### 2020-01-22 Inversion_results.Rmd
Many major changes. 1. Best-fit param changed so that not just top 10% of best-fits for each par.sam

---
title: "BCI Water Uptake Depths"
author: "Rutuja Chitra-Tarak"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Rooting profiles used for inversion

```{r Chunk 0.01, eval = TRUE, echo = FALSE, message = FALSE, fig.align = 'center', out.width = "100%", fig.pos = 'H', fig.cap = ""}
knitr::include_graphics(file.path("figures/rooting_profiles_for_inversion.jpeg"))
```
## All size-classes

```{r Chunk 2.00, eval = TRUE, echo = FALSE, message = FALSE, fig.align = 'center', out.width = "100%", fig.pos = 'H', fig.cap = ""}
knitr::include_graphics(file.path("figures/UDI/mean/sp_uptake.depth_all_sizes_cor0.3_mean_skIqAp.jpeg"))
knitr::include_graphics(file.path("figures/UDI/mean/violinplot_all_sizes_cor0.3_mean_skIqAp.jpeg"))
```

### Species Uptake Depth Index versus delta Deuterium in Xylem sap water

#### *New Model*
```{r Chunk 0.15, eval = TRUE, echo = FALSE, message = FALSE, fig.align = 'center', out.width = "100%", fig.pos = 'H', fig.cap = ""}
knitr::include_graphics(file.path("figures/UDI_confidence/mean/Comparison_with_Meinzer1999_deltaD_vs.modelled_uptake.depth_cor0.3_comm_with outliers.jpeg"))
knitr::include_graphics(file.path("figures/UDI_confidence/mean/Comparison_with_Meinzer1999_deltaD_vs.modelled_uptake.depth_cor0.3_comm_without outliers.jpeg"))
knitr::include_graphics(file.path("figures/UDI_confidence/mean/Comparison_with_Meinzer1999_deltaD_vs.modelled_uptake.depth_cor0.3_sp_with outliers.jpeg"))
knitr::include_graphics(file.path("figures/UDI_confidence/mean/Comparison_with_Meinzer1999_deltaD_vs.modelled_uptake.depth_cor0.3_sp_without outliers.jpeg"))
```
#### *Old Model*

```{r Chunk 0.16, eval = TRUE, echo = FALSE, message = FALSE, fig.align = 'center', out.width = "100%", fig.pos = 'H', fig.cap = ""}
knitr::include_graphics(file.path("figures/UDI_confidence/mean/Old_model_Comparison_with_Meizer1999_isotope_vs.modelled_uptake.depth.tiff"))
```

#### Meinzer et al. 1999: Relation between delta Deuterium in soil water versus Depth

```{r Chunk 0.17, eval = TRUE, echo = FALSE, message = FALSE, fig.align = 'center', out.width = "100%", fig.pos = 'H', fig.cap = ""}
knitr::include_graphics(file.path("figures/UDI_confidence/Meinzer_etal_1999_Depth_vs_soil_deltaD.jpeg"))
```

#### New Model: Water-Uptake Depth versus Meinzer et al 1999 based water-uptake depth

```{r Chunk 0.18, eval = TRUE, echo = FALSE, message = FALSE, fig.align = 'center', out.width = "100%", fig.pos = 'H', fig.cap = ""}
knitr::include_graphics(file.path("figures/UDI_confidence/mean/Comparison_with_Meinzer1999_isotopic_vs.modelled_uptake.depth_cor0.3_sp_without outliers.jpeg"))
```

### Species-size growth by mean interval Growth Factor

```{r Chunk 0.10, eval = TRUE, echo = FALSE, message = FALSE, fig.align = 'center', out.width = "100%", fig.pos = 'H', fig.cap = ""}
knitr::include_graphics(file.path("figures/mean_growth_by_gfac_by_size.jpeg"))
```

### Species-size growth by mean interval Solar Radiation

```{r Chunk 0.101, eval = TRUE, echo = FALSE, message = FALSE, fig.align = 'center', out.width = "100%", fig.pos = 'H', fig.cap = ""}

knitr::include_graphics(file.path("figures/mean_growth_by_solar_radiation_by_size.jpeg"))
```

### Mean Rsq between
```{r Chunk 0.11, eval = FALSE, echo = FALSE, message = FALSE, fig.align = 'center', out.width = "100%", fig.pos = 'H', fig.cap = ""}

knitr::include_graphics(file.path("figures/UDI/mean/UDI_mean.rsq_size_class_cor0.3_2019-10-14_5000_mean_5_rsq_vs_UDI.jpeg"))
```

### Does a species' water-Uptake Depth Index explain its mean mortality rate?

#### Mortality continues after the 1982-83 El-Nino drought

```{r Chunk 3.00, eval = TRUE, echo = FALSE, message = FALSE, fig.align = 'center', out.width = "100%", fig.pos = 'H', fig.cap = ""}

knitr::include_graphics(file.path("figures/mortality/sp_Mortality_rate_by_period_above1cmDBH_aboveN50.jpeg"))
```

##### Species mean abundance cutoff n = 1
```{r Chunk 3.021, eval = TRUE, echo = FALSE, message = FALSE, fig.align = 'center', out.width = "100%", fig.pos = 'H', fig.cap = ""}
knitr::include_graphics(file.path("figures/mortality/mean/sp_Mortality_rate_by_period_udi_by_size_aboveN1_tlpcommn.jpeg"))
knitr::include_graphics(file.path("figures/mortality/mean/sp_Mortality_rate_by_period_udi_by_size_aboveN1_tlpsp.jpeg"))
```

##### Species mean abundance cutoff n = 10
```{r Chunk 3.022, eval = TRUE, echo = FALSE, message = FALSE, fig.align = 'center', out.width = "100%", fig.pos = 'H', fig.cap = ""}
knitr::include_graphics(file.path("figures/mortality/mean/sp_Mortality_rate_by_period_udi_by_size_aboveN10_tlpcommn.jpeg"))
knitr::include_graphics(file.path("figures/mortality/mean/sp_Mortality_rate_by_period_udi_by_size_aboveN10_tlpsp.jpeg"))
```

##### Species mean abundance cutoff n = 50
```{r Chunk 3.023, eval = TRUE, echo = FALSE, message = FALSE, fig.align = 'center', out.width = "100%", fig.pos = 'H', fig.cap = ""}
knitr::include_graphics(file.path("figures/mortality/mean/sp_Mortality_rate_by_period_udi_by_size_aboveN50_tlpcommn.jpeg"))
knitr::include_graphics(file.path("figures/mortality/mean/sp_Mortality_rate_by_period_udi_by_size_aboveN50_tlpsp.jpeg"))
```

### Species Water-Uptake Depth Index is not correlated with its mean growth rate.

```{r Chunk 3.031, eval = TRUE, echo = FALSE, message = FALSE, fig.align = 'center', out.width = "100%", fig.pos = 'H', fig.cap = ""}
knitr::include_graphics(file.path("figures/mortality/mean/sp_mean_Growth_vs_udi_size_tlpcomm.jpeg"))
```

### Species growth vs mortality rate

```{r Chunk 3.04, eval = TRUE, echo = FALSE, message = FALSE, fig.align = 'center', out.width = "100%", fig.pos = 'H', fig.cap = ""}
knitr::include_graphics(file.path("figures/mortality/mean/sp_mean_Growth_vs_mrate.jpeg"))
```

### Is Water-Uptake Depth Index related to other hydraulic traits?
#### TLP at Common/Community level
#### TLPlevel = comm, *Tiny*
```{r Chunk 4.01, eval = TRUE, echo = FALSE, message = FALSE, fig.align = 'center', out.width = "100%", fig.pos = 'H', fig.cap = ""}
knitr::include_graphics(file.path("figures/traits/mean/pca/sp_UDI_traits_democomm_tiny.tiff"))
```
#### TLPlevel = comm, *Small*
```{r Chunk 4.02, eval = TRUE, echo = FALSE, message = FALSE, fig.align = 'center', out.width = "100%", fig.pos = 'H', fig.cap = ""}
knitr::include_graphics(file.path("figures/traits/mean/pca/sp_UDI_traits_democomm_small.tiff"))
```
#### TLPlevel = comm, *Medium*
```{r Chunk 4.03, eval = TRUE, echo = FALSE, message = FALSE, fig.align = 'center', out.width = "100%", fig.pos = 'H', fig.cap = ""}
knitr::include_graphics(file.path("figures/traits/mean/pca/sp_UDI_traits_democomm_medium.tiff"))
```
#### TLPlevel = comm, *Large*
```{r Chunk 4.04, eval = TRUE, echo = FALSE, message = FALSE, fig.align = 'center', out.width = "100%", fig.pos = 'H', fig.cap = ""}
knitr::include_graphics(file.path("figures/traits/mean/pca/sp_UDI_traits_democomm_large.tiff"))
```

#### TLP at Species level
#### TLPlevel = sp, *Tiny*
```{r Chunk 5.01, eval = TRUE, echo = FALSE, message = FALSE, fig.align = 'center', out.width = "100%", fig.pos = 'H', fig.cap = ""}
knitr::include_graphics(file.path("figures/traits/mean/pca/sp_UDI_traits_demosp_tiny.tiff"))
```
#### TLPlevel = sp, *Small*
```{r Chunk 5.02, eval = TRUE, echo = FALSE, message = FALSE, fig.align = 'center', out.width = "100%", fig.pos = 'H', fig.cap = ""}
knitr::include_graphics(file.path("figures/traits/mean/pca/sp_UDI_traits_demosp_small.tiff"))
```
#### TLPlevel = sp, *Medium*
```{r Chunk 5.03, eval = TRUE, echo = FALSE, message = FALSE, fig.align = 'center', out.width = "100%", fig.pos = 'H', fig.cap = ""}
knitr::include_graphics(file.path("figures/traits/mean/pca/sp_UDI_traits_demosp_medium.tiff"))
```
#### TLPlevel = sp, *Large*
```{r Chunk 5.04, eval = TRUE, echo = FALSE, message = FALSE, fig.align = 'center', out.width = "100%", fig.pos = 'H', fig.cap = ""}
knitr::include_graphics(file.path("figures/traits/mean/pca/sp_UDI_traits_demosp_large.tiff"))
```

#### TLP at Common/Community level
```{r Chunk 4.11, eval = TRUE, echo = FALSE, message = FALSE, fig.align = 'center', out.width = "100%", fig.pos = 'H', fig.cap = ""}
knitr::include_graphics(file.path("figures/traits/mean/sp_UDI_vs_traits_corrplot_tlpcomm_tiny_demo.jpeg"))
knitr::include_graphics(file.path("figures/traits/mean/sp_UDI_vs_traits_corrplot_tlpcomm_small_demo.jpeg"))
knitr::include_graphics(file.path("figures/traits/mean/sp_UDI_vs_traits_corrplot_tlpcomm_medium_demo.jpeg"))
knitr::include_graphics(file.path("figures/traits/mean/sp_UDI_vs_traits_corrplot_tlpcomm_large_demo.jpeg"))
```

#### TLP at Species level
```{r Chunk 5.11, eval = TRUE, echo = FALSE, message = FALSE, fig.align = 'center', out.width = "100%", fig.pos = 'H', fig.cap = ""}
knitr::include_graphics(file.path("figures/traits/mean/sp_UDI_vs_traits_corrplot_tlpsp_tiny_demo.jpeg"))
knitr::include_graphics(file.path("figures/traits/mean/sp_UDI_vs_traits_corrplot_tlpsp_small_demo.jpeg"))
knitr::include_graphics(file.path("figures/traits/mean/sp_UDI_vs_traits_corrplot_tlpsp_medium_demo.jpeg"))
knitr::include_graphics(file.path("figures/traits/mean/sp_UDI_vs_traits_corrplot_tlpsp_large_demo.jpeg"))
```