### 2020-04-11 Workflow.Rmd
added likelihood calculcation

---
title: "Workflow of Inverse Model for Rooting Profiles"
author: "Rutuja"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, include = TRUE)
```

## How is this project set-up

### 1.GLUEsetup_part1_BCI.R

 * Loads best.fit ensembles of Soil Water Potential (SWP) from bci.elm.fates.hydro::psi of a specific version (given by the "current folder")
 * Loads hypothetical rooting profiles generated by R/3.090_root_parameters_b.R
 * Loads Turgor Loss Point (TLP) data
 * Loads tree growth census dates for specific no. of intervals
 * Calculates btran time-series by census-interval for each root profile for all the best.fit SWP ensemble
  + by individual species with species-specific TLP (N of btran time series == No. of species x No. of SWP ensembles x No. of root profiles)
  + at community-mean TLP (No. of btran time series == No. of SWP ensembles x No. of root profiles)
  * Saves all the above info in a list called info, except species level btran is saved in info.2

## Then using a combination of a range of hypothetical root profiles obtained from 2.rooting_profiles.R and SWP ensemble simulations generate a btran time-series using equation : Btran= i=1 to i=z RootFrac_z * f(swp_z),
where f(swp_z) is a growth.factor_z based on a linear step relationship with swp_z ap species specific or community mean TLP level.
Code/6.0_Growth_correlate_table_from_swp.R

```{r Chunk 1.0, eval = FALSE}
source("code/4.1.1.GLUEsetup_part1_BCI.R")
GLUEsetup_part1(current.folder = "2019-10-14_5000", intervals = 5) # this can take 30 min to get btran by nsam
```

```{r Chunk 2, eval = FALSE}
source("code/4.1.2.GLUEsetup_part2_BCI.R")
GLUEsetup_part2(growth.type = "med", dbh.residuals = "on", solar.residuals = "off", 
                growth.selection = "size_class_predefined_cc_scaled") #"size_class_varying_non_cc") #n.ensembles = 100000
# growth.type = "med"; dbh.residuals = "on"; solar.residuals = "off"; growth.selection = "size_class_predefined_cc_scaled"
```

```{r Chunk 3, eval = TRUE}
# This sources "code/4.2GLUErun_part2_func_BCI.R"
source("code/4.3.GLUErun_part2_BCI.R")
# splevel = "off"; goodness.fit = 0.3; drop.months = "Mar"
GLUE_run(splevel = "off", goodness.fit = 0.3, drop.months = "None")
GLUE_run(splevel = "off", goodness.fit = 0.3, drop.months = "Jan")
GLUE_run(splevel = "off", goodness.fit = 0.3, drop.months = "Feb")
GLUE_run(splevel = "off", goodness.fit = 0.3, drop.months = "Mar")
GLUE_run(splevel = "off", goodness.fit = 0.3, drop.months = "JanFeb")
GLUE_run(splevel = "off", goodness.fit = 0.3, drop.months = "FebMar")
GLUE_run(splevel = "off", goodness.fit = 0.3, drop.months = "JanFebMar")
GLUE_run(splevel = "on", goodness.fit = 0.3, drop.months = "None")
GLUE_run(splevel = "on", goodness.fit = 0.3, drop.months = "Jan")
GLUE_run(splevel = "on", goodness.fit = 0.3, drop.months = "Feb")
GLUE_run(splevel = "on", goodness.fit = 0.3, drop.months = "Mar")
GLUE_run(splevel = "on", goodness.fit = 0.3, drop.months = "JanFeb")
GLUE_run(splevel = "on", goodness.fit = 0.3, drop.months = "FebMar")
GLUE_run(splevel = "on", goodness.fit = 0.3, drop.months = "JanFebMar")
```

<!-- ```{r Chunk 2.4.0, eval = TRUE} -->
<!-- source("code/4.4.Best-fit_param_BCI.R") -->
<!-- # bestfit.param(splevel = "on") -->
<!-- bestfit.param(splevel = "off") -->
<!-- ``` -->

```{r Chunk 4, eval = TRUE}
source("code/4.5.prep.udi.calculator.R")
# splevel = "on"; drop.months = "JanFebMar"
prep.udi.calculator(splevel = "on", drop.months = "Feb")
prep.udi.calculator(splevel = "on", drop.months = "FebMar")
prep.udi.calculator(splevel = "on", drop.months = "Mar")

prep.udi.calculator(splevel = "off", drop.months = "Feb")
prep.udi.calculator(splevel = "off", drop.months = "FebMar")
prep.udi.calculator(splevel = "off", drop.months = "Mar")
```

```{r Chunk 5, eval = TRUE}
source("code/4.6.udi.calculator.R")
##---needs to be run manually on macbook for splevel = "off" and iso.subset = "off"---
# udi.calculator(splevel = "on", dryseason = "on", rsq.thresh = 0.5, iso.subset = "on") ## udi is not calculated below this rsq.thresh
# splevel = "on"; dryseason = "on"; rsq.thresh = 0.5;  n.ci.thresh = 3; iso.subset = "on"; drop.months = "JanFebMar"
udi.calculator(splevel = "on", dryseason = "on", rsq.thresh = 0.3, 
               n.ci.thresh = 3, iso.subset = "on", drop.months = "Feb")
udi.calculator(splevel = "on", dryseason = "on", rsq.thresh = 0.3, 
               n.ci.thresh = 3, iso.subset = "on", drop.months = "FebMar")
udi.calculator(splevel = "on", dryseason = "on", rsq.thresh = 0.3, 
               n.ci.thresh = 3, iso.subset = "on", drop.months = "Mar")

udi.calculator(splevel = "off", dryseason = "on", rsq.thresh = 0.3, 
               n.ci.thresh = 3, iso.subset = "on", drop.months = "Feb")
udi.calculator(splevel = "off", dryseason = "on", rsq.thresh = 0.3, 
               n.ci.thresh = 3, iso.subset = "on", drop.months = "FebMar")
udi.calculator(splevel = "off", dryseason = "on", rsq.thresh = 0.3, 
               n.ci.thresh = 3, iso.subset = "on", drop.months = "Mar")

udi.calculator(splevel = "on", dryseason = "on", rsq.thresh = 0.3, 
               n.ci.thresh = 3, iso.subset = "off", drop.months = "Feb")
udi.calculator(splevel = "on", dryseason = "on", rsq.thresh = 0.3, 
               n.ci.thresh = 3, iso.subset = "off", drop.months = "FebMar")
udi.calculator(splevel = "on", dryseason = "on", rsq.thresh = 0.3, 
               n.ci.thresh = 3, iso.subset = "off", drop.months = "Mar")

udi.calculator(splevel = "off", dryseason = "on", rsq.thresh = 0.3, 
               n.ci.thresh = 3, iso.subset = "off", drop.months = "Feb")
udi.calculator(splevel = "off", dryseason = "on", rsq.thresh = 0.3, 
               n.ci.thresh = 3, iso.subset = "off", drop.months = "FebMar")
udi.calculator(splevel = "off", dryseason = "on", rsq.thresh = 0.3, 
               n.ci.thresh = 3, iso.subset = "off", drop.months = "Mar")
```

```{r Chunk 6, eval = TRUE}
source("code/4.7.summarising_udi_by_sp_size.R")
summarise.udi(splevel = "on", dryseason = "on", rsq.thresh = 0.5, root.selection = "exp", iso.subset = "on", drop.months = "Feb")
summarise.udi(splevel = "on", dryseason = "on", rsq.thresh = 0.5, root.selection = "exp", iso.subset = "on", drop.months = "FebMar")
# # splevel = "on"; dryseason = "off"; rsq.thresh = 0.5; root.selection = "exp"; iso.subset = "off"
summarise.udi(splevel = "on", dryseason = "on", rsq.thresh = 0.5, root.selection = "exp", iso.subset = "on", drop.months = "Mar")

summarise.udi(splevel = "off", dryseason = "on", rsq.thresh = 0.5, root.selection = "exp", iso.subset = "on", drop.months = "FebMar")
summarise.udi(splevel = "off", dryseason = "on", rsq.thresh = 0.5, root.selection = "exp", iso.subset = "on", drop.months = "Feb")
summarise.udi(splevel = "off", dryseason = "on", rsq.thresh = 0.5, root.selection = "exp", iso.subset = "on", drop.months = "Mar")

summarise.udi(splevel = "on", dryseason = "on", rsq.thresh = 0.5, root.selection = "exp", iso.subset = "off", drop.months = "Feb")
summarise.udi(splevel = "on", dryseason = "on", rsq.thresh = 0.5, root.selection = "exp", iso.subset = "off", drop.months = "FebMar")
summarise.udi(splevel = "on", dryseason = "on", rsq.thresh = 0.5, root.selection = "exp", iso.subset = "off", drop.months = "Mar")

summarise.udi(splevel = "off", dryseason = "on", rsq.thresh = 0.5, root.selection = "exp", iso.subset = "off", drop.months = "FebMar")
summarise.udi(splevel = "off", dryseason = "on", rsq.thresh = 0.5, root.selection = "exp", iso.subset = "off", drop.months = "Feb")
summarise.udi(splevel = "off", dryseason = "on", rsq.thresh = 0.5, root.selection = "exp", iso.subset = "off", drop.months = "Mar")
```

```{r Chunk 7, eval = TRUE}
source("code/10.Comparison_with_Meinzer1999.R")
iso.compare(goodness.fit = 0.3, dryseason = "on", 
            root.selection = "exp", iso.subset = "on", drop.months = "None")
iso.compare(goodness.fit = 0.3, dryseason = "on", 
            root.selection = "exp", iso.subset = "on", drop.months = "JanFebMar")
iso.compare(goodness.fit = 0.3, dryseason = "on", 
            root.selection = "exp", iso.subset = "on", drop.months = "JanFeb")
iso.compare(goodness.fit = 0.3, dryseason = "on", 
            root.selection = "exp", iso.subset = "on", drop.months = "Jan")
# goodness.fit = 0.3; dryseason = "on"; root.selection = "exp"; iso.subset = "on"; drop.months = "None"
```

```{r Chunk 8, eval = FALSE}
source("code/9.6_mortality_BCI.R")
demo.graphs(level.folder = "commlevel", n.threshold = 30, dbh.residuals = "on", 
           dryseason = "on", root.selection = "on", iso.subset = "off")
# level.folder = "commlevel"; n.threshold = 30; dbh.residuals = "on";
#            dryseason = "on"; root.selection = "on"; iso.subset = "off"
```

```{r Chunk 9, eval = FALSE}
source("code/5.species_graphs_BCI.R")
source("code/13.hydraulic_traits_BCI.R")
```