### 2020-10-17 Appendix.Rmd
First commit

---
title: "Drought-sensitive trees survive on deep water access in a tropical forest community"
subtitle: 'Appendix'
author: ''
date: ''
output:
  bookdown::html_document2:
    bookdown::word_document2:
        reference_docx: manuscriptstyle.docx
        fig_caption: yes
  number_sections: no
theme: flatly
bookdown::pdf_document2: default
csl: new-phytologist.csl
bibliography: references.bib
number_sections: true
# header-includes: 
#   \usepackage{float} \floatplacement{figure}{H} 
#   \setcounter{table}{0}  
#   \renewcommand{\thetable}{S\arabic{table}} 
#   \setcounter{figure}{0} 
#   \renewcommand{\thefigure}{S\arabic{figure}}
---

```{r, include=FALSE, echo = FALSE}
if(!require("pacman", quietly=TRUE)) install.packages("pacman")
suppressPackageStartupMessages(
  pacman::p_load(bookdown, knitr, pander, citr, broom))
options(tinytex.verbose = TRUE)
source("code/load.R")
source("code/manuscript_data_figs.R")
# load("results/manuWorkSpace.RData")
# install.packages(c("vctrs", "rlang"))
```

```{r setup, echo = FALSE}
library(knitr)
opts_knit$set(eval.after = "fig.cap")
```

\documentclass{article}

\usepackage{chngcntr}
\begin{document}

\section{Introduction}

\subsection{Problem description}

\begin{figure}[htp]
  \centering Figure
  \caption{Call this figure 1.}
  \label{fig1}
\end{figure}

S\@ref(fig:other-erd-mods)). 
S\@ref(fig:erd-mort-census)). 
\appendix
\counterwithin{figure}{section}

\section{Figures}

\beginsupplement

```{r other-erd-mods, eval = TRUE, echo = FALSE, message = FALSE, fig.align = 'center', out.width = "80%", fig.pos = 'H', results = "asis", fig.cap = paste0("\\label{lab-other-erd-mods}Effective Rooting Depth versus $\\delta^2H$ for each of the ERD models tested (panels). For each ERD model only those species have an estimate of ERD that have passed the test of Pearson's $\\rho$ > 0 and *R*^2^ > 0.5 during ERD parameterisation. The number of species with an ERD estimate as well as $\\delta^2H$. or each model are shown in inset, along with the goodness-of-fit (*R*^2^) and *P*-values for a linear relationship between ERD and $\\delta^2H$. See Eqs. \\@ref(eq:M1)-\\@ref(eq:M6). See Table \\@ref(tab:ab-param) for the full form of scientific names.")}
knitr::include_graphics(file.path("figures/PhenoDemoTraitsPsi/psi.corr_best.depth_xylem_sap_deltaD_sp_color_Meinzer.jpeg"))
```

```{r erd-mort-census, eval = TRUE, echo = FALSE, message = FALSE, fig.align ='center', out.width = "100%", fig.pos = 'H', results = "asis", fig.cap = "\\label{lab-erd-mort-census}Species mortality rate in each census interval during the study period plotted against ERD."}
knitr::include_graphics(file.path("figures/PhenoDemoTraitsPsi/mortality_by rdi.gr.jpeg"))
```

\end{document}