### 2020-10-24 old_ms_sections.Rmd
# moved in discussion on Chitra-Tarak 2018

---
title: "Hydraulically-vulnerable trees are buffered by deep-water access in a tropical forest community"
author: ''
date: ''
output:
  bookdown::html_document2:
    number_sections: no
    theme: flatly
    fig_caption: yes
  bookdown::word_document2:
        reference_docx: manuscriptstyle.docx
        fig_caption: yes
  bookdown::pdf_document2: default
bibliography: references.bib
csl: new-phytologist.csl
---

```{r, include=FALSE, echo = FALSE}
if(!require("pacman", quietly=TRUE)) install.packages("pacman")
suppressPackageStartupMessages(
  pacman::p_load(bookdown, knitr, pander, citr, broom))
options(tinytex.verbose = TRUE)
source("code/load.R")
source("code/manuscript_data_figs.R")
```

```{r setup, echo = FALSE}
library(knitr)
opts_knit$set(eval.after = "fig.cap")
```

## Intro

Historic climatic variation is relatively small (Mora et al., 2013), and, though data have been scant, trees' hydraulic safety margins are reportedly relatively narrow (Choat et al., 2012; but see Oliveira et al., 2019).

## M&M

For each species $s$ we calculated mean mortality rate across all intervals as $\textrm{Mort}_{s,\textrm{mean}} = \frac{1}{6}\sum_{t=1}^{6} {\textrm{Mort}_{s,t}}$.

## Results

The hydraulic efficiency-safety trade-off between high values of maximum stem conductivity (high $K_\textrm{max,stem}$) and cavitation-resistance (low $\Psi$~50,stem~ or $\Psi$~88,stem~) is hypothesized although community-wide examples are rare [@Hacke:2006; @Gleason.2016]. We report such a trade-off ($n$ = 26 species) at our study site stem level.

Species with deeper ERD showed lower mean mortality rates across the 35-yrs record (1982-2015; *R*^2^ = 0.11, *p* = 0.075, *n* = `r length(erd.mrate.sp)`; Fig. \@ref(fig:erd-mort)), and this relationship was stronger among evergreen species.

Specifically for the water environment (or, resource), observations of individual trade-offs between environment-trait, trait-trait and trait-demography has identified an acquisitive-conservative resource economic spectrum similar to those for light and nutrients, although, for water, examples of all trade-offs observed simultaneously are rare, and below-ground traits or environment linkages are weak [@Reich:2014; @Sterck:2011]. Acquisitive species have traits with high capacity to acquire, move, store and use water (higher hydraulic conductivity, low wood density) and achieve high growth, but build flimsy tissues (high specific leaf area) and are less tolerant of low resource availablity, whereas species with conservative traits are better protected from carbon losses and are more physiologically drought tolerant (greater embolism resistance, larger safety margins). Whether and how the acquisitive-conservative spectrum holds true for the water dynamics vertically below-ground is not clear. Acquisitive species are spatially associated with wet sites (high resource) such as valleys or those with high rainfall [@Kursar:2009; @Oliveira:2019]. Deep rooting may be expected to allow an acquisitive strategy by enabling a steady water supply. However, the degree to which deep rooting is associated with an acquisitive strategy may depend on the site-specific dynamics of water availability by depth, which is driven by precipitation regime and hydrology. For example, in pulse-driven systems *shallow*-rooted plants are associated with acquisitive traits (higher hydraulic conductivity) and opportunistic strategies (e.g. dry deciduousness) that maximise the use of the water-pulse, while *deep*-rooted plants show conservative traits (cavitation-resistant xylem, osmoregulation) [@Schwinning:2013 and references therein].

\
At the community level, we filled data gaps in $LMA$ for `r sum(c.lma.gaps[-which(names(c.lma.gaps) == "original")])` species (of total `r sum(c.lma.gaps)`)  using linear relationships with other variables: `r c.lma.gaps[[which(names(c.lma.gaps) == "filled-LMA-DISC")]]` species with $LMA_{disc}$, `r c.lma.gaps[[which(names(c.lma.gaps) == "filled-LMA-Nobby")]]` with $LMA_{lamina}$, and `r c.lma.gaps[[which(names(c.lma.gaps) == "filled-LEAF")]]` with $LMA_{leaf}$, defined as the mean leaf mass per unit area measured for the entire leaf including the petiole (g m^-2^) for leaves receiving direct sunlight ($\overline{LMA} = `r leaf.cf[1]` + `r leaf.cf[2]`LMA_{leaf}$; Adj. *R*^2^ = `r leaf.r2`; *P* `r leaf.p`, *n* = `r leaf.n`). For the `r sum(c.lma.gaps)` species thus with $LMA$, we gap-filled data on  WSG for `r sum(c.wsg.gaps[-which(names(c.wsg.gaps) == "original")])` species using a linear relationship with wood specific gravity reported by @Chave:2009 ($\overline{WSG} = `r chave.cf[1]` + `r chave.cf[2]`WSG_{chave}$; Adj. *R*^2^ = `r chave.r2`; *P* `r chave.p`, *n* = `r chave.n`).

From Discussion first para

Mortality rates of shallow ERD species may have been exacerbated by increased drought events during the study period. @Condit.2017 showed that half of the extreme dry seasons (< 1 SD annual water deficit) during 1931-2015 (85-yr) occurred during 1982-2005 (24-yr) of the 34-yr long study period (1982-2015).

The model in @Chitra-Tarak.2018 estimated six species-specific phenomenological parameters: three as root fraction for corresponding lumped vadose zone layers, and three as shape parameters for water uptake vs availability curves. For each species, they calculated daily water uptake successively from three layers, that is, the supply, to satisfy the demand of the daily maximum evapotranspiration, a function of PET and LAI. They regressed total supply to total demand ratio over census intervals as a growth coefficient against growth (five data points), and chose values for six parameters that best-fit (highest *R*^2^, that is, two additional parameters). Our inverse ERD model fits only two parameters by way of regressing depth-specific model against five growth data points. Our model was constrained on VPD and species leaf vulnerability curves, variables that are shown to be physiologically meaningful.

The model in CT2018 were not corroborated with such data as existing methods were inadequate to extract water bound to the saprolite found in their dry forest study site (Jean Riotte, pers. comm.). The model of below-ground water-availability in this study (ELM-FATES) differs from the conceptual water balance model used by CT2018 and was calibrated with data on long-term soil moisture dynamics at a point location (0-0.15m) and plot-wide measurements by depth (0.15, 0.4 and 1 m), as well as, evapotranspiration and stream discharge. The model in CT2018 was calibrated only on stream discharge and water table levels.