### 2020-11-06 manuscript.rmd
1. edited discussion

---
title: "Hydraulically-vulnerable trees survive on deep-water access during droughts in a tropical forest"
author: ''
date: ''
output:
  bookdown::word_document2:
        reference_docx: manuscriptstyle.docx
        fig_caption: yes
  bookdown::pdf_document2: default
  bookdown::html_document2:
    number_sections: no
    theme: flatly
    fig_caption: yes
bibliography: references.bib
csl: new-phytologist.csl
---

```{r, include=FALSE, echo = FALSE}
if(!require("pacman", quietly=TRUE)) install.packages("pacman")
suppressPackageStartupMessages(
  pacman::p_load(bookdown, knitr, pander, citr, broom))
options(tinytex.verbose = TRUE)
source("code/load.R")
source("code/manuscript_data_figs.R")
# load("results/manuWorkSpace.RData")
# install.packages(c("vctrs", "rlang"))
```

```{r setup, echo = FALSE}
library(knitr)
opts_knit$set(eval.after = "fig.cap")
```

Chitra-Tarak, Rutuja*^1,2^, Chonggang Xu^1^, Salomón Aguilar^3^, Kristina Anderson-Teixeira^3,4^, Jeff Chambers^5^, Matteo Detto^3,6^, Boris Faybishenko^5^, Rosie A. Fisher^7^, Ryan Knox^5^, Charlie Koven^5^, Lara Kueppers^5,8^, Nobert Kunert^3,4,9^, Stefan J. Kupers^10^, Nathan G. McDowell^11,12^, Brent D. Newman^1^, Steven R. Paton^3^, Rolando Pérez^3^, Laurent Ruiz^13,14,15^, Lawren Sack^16^, Jeffrey M. Warren^17^, Brett T. Wolfe^3,18^, Cynthia Wright^17^, S. Joseph Wright^3^, Joseph Zailaa^4,16,19^, Sean M. McMahon^2,3^

\
*__Corresponding author__: Rutuja Chitra-Tarak  
E: arutuj@gmail.com ;  T: +1 (443) 758 1257

1. Los Alamos National Laboratory, Earth and Environmental Sciences Division (EES-14) 
MS J495, Los Alamos, NM, USA 87545-1663 
2. Smithsonian Environmental Research Center, 647 Contees Wharf Road, Edgewater, MD, USA 21037-0028 
3. Smithsonian Tropical Research Institute, Panama, PA 
4. Conservation Ecology Center, Smithsonian Conservation Biology Institute, Front Royal, VA, USA 
5. Lawrence Berkeley National Laboratory, Climate and Ecosystem Sciences Division, Berkeley, CA, USA 
6. Princeton University, Ecology and Evolutionary Biology, Princeton, NJ, USA 
7. National Center for Atmospheric Research, Climate and Global Dynamics Division
Boulder, CO, USA 
8. University of California Berkeley, Energy and Resources Group, 310 Barrows Hall #3050
Berkeley, CA, USA 94720 
9. Institute of Botany, University of Natural Resources and Life Sciences Vienna, Department of Integrative Biology and Biodiversity Research, Gregor‐Mendel‐Str 33, A‐1190 Wien, Wien, AT 
10. German Centre for Integrative Biodiversity Research (iDiv) Halle-Jena-Leipzig, Computational Forest Ecology, Leipzig, Saxony, DE 
11. Atmospheric Sciences and Global Change Division, Pacific Northwest National Lab, PO Box 999, Richland, WA, USA 99352 
12. School of Biological Sciences, Washington State University, PO Box 644236, Pullman, WA, USA 99164-4236 
13. Indo-French Cell for Water Sciences, Indian Institute of Science, Bangalore, Karnataka, IN 
14. IRD, CNRS, UPS, UMR GET, Toulouse, FR 
15. INRAE, Institut Agro, UMR SAS, Rennes, FR 
16. University of California Los Angeles, Ecology and Evolutionary Biology, 612 Charles E. Young Drive South, Los Angeles, CA, USA 90095 
17. Oak Ridge National Laboratory, Environmental Sciences Division, Oak Ridge, TN, USA 
18. School of Renewable Natural Resources, Louisiana State University Agricultural Center, Baton Rouge, LA 70803, USA 
19. California State University Los Angeles, Biological Sciences Department, Los Angeles, CA, USA

\

Section  | Word Count  
------------- | -------------  
Summary | 197  
Main Text  | 6487  
Introduction | 832  
Methods | 3277  
Results | 876  
Discussion | 1502  
Keywords | 6

\

Item | Count  
------------- | -------------  
Figures in color | 9  
Tables | 2  
Supplementary Material |  1 file

# Summary

- Deep-water access is arguably the most long-standing, but under-studied mechanisms that plants employ to survive during drought. Vulnerability to embolism and hydraulic safety margins can predict mortality risk at given levels of dehydration, but deep-water access may delay plant dehydration. Here, we tested the role of deep-water access in enabling survival within a diverse tropical forest community in Panama using a novel data-model approach.
- We inversely estimated the effective rooting depth (ERD, as the average depth of water extraction), for 32 canopy species by linking diameter growth dynamics (1990-2015) to vapor pressure deficit, water potentials in the whole-soil column, and leaf hydraulic vulnerability curves. We validated ERD estimates against existing empirical data of potential water access depths.
- Across species, deeper ERD was associated with higher maximum stem hydraulic conductivity, greater vulnerability to xylem embolism, narrower safety margins, and lower mortality rates during extreme droughts over 35 years (1981-2015). Species with deeper ERD did not cross critical hydraulic thresholds indicating that trees compensate for water-stress related mortality risk through deep-water access.
- The role of deep-water access in mitigating mortality of hydraulically-vulnerable trees has important implications for our predictive understanding of forest dynamics under current and future climates.
\
Keywords: rooting depths, deep-water access, hydraulic vulnerability and safety margins, drought-induced mortality, drought tolerance, tropical forest, safety-efficiency trade-off, hydrological droughts

# Introduction

Drought-induced mortality in tropical forests may have significant global implications. Tropical forests play a disproportionately large role in the global carbon and energy cycles [@Bonan:2008], and support half of global biodiversity [@Wright:20050th], but face a threat from intensifying droughts [@Doughty:2015; @Xu.2019]. Tropical forests have been expected to be an especially drought-vulnerable biome given the combination of climate risk and vegetation sensitivity [@Meir:2015trw]. Still, mortality events in the tropics are rarely as large as those in temperate and boreal zones [@McDowell.2018], leading to questions regarding the role of large trait-diversity and hydraulic strategies in mitigating mortality events. Nevertheless, mortality rates are increasing in some tropical regions [@Brienen.2015; @Hubau.2020] and widespread drought-induced tree mortality has occurred across the tropics for specific functional groups [@Bennett:2015; @Chitra-Tarak.2018; @Hilker:2014; @Phillips:2010]. State-of-the-art dynamic global vegetation models (DGVMs) struggle to capture these drought-induced vegetation dynamics in the tropics [@Galbraith.2010; @Powell:2013; @Powell.2018], because underlying mechanisms of drought-tolerance are not fully understood nor quantified.

\
Plants rely on a variety of structural and functional mechanisms to avoid or tolerate a drought, from adjustment of leaf area (deciduousness), change in leaf angle, reductions in stomatal conductance, upregulation of aquaporins, osmotic regulation, stem water storage, deep-water access, increased root production, hydraulic redistribution and embolism resistance [@McDowell.2008]. Deep-water access is arguably the most long-standing, yet under-studied mechanism. As plant-available water varies with depth, trees in the same forest with different rooting depths, depending on species and size [@Meinzer:1999; @Chitra-Tarak.2018; @Brum.2019], differ in their experience during a drought, and thus in their growth and mortality responses [@Chitra-Tarak.2018]. A key bottleneck in community-wide testing of this mechanism has been a lack of data in both trees' rooting or water-sourcing depths and plant-available soil water at those depths. A recent meta-analysis documented maximum rooting depths for 318 tree species, that is, less than 0.5% of the world’s more than 60 thousand tree species [@Fan:2017; @Beech:2017]. Further, only a small fraction of those are tropical, even though more than 90% of the world’s tree diversity resides in the tropics [@Slik:2015]. The use of stable isotopes of water as a tracer provides an indirect measure of water sourcing depths by matching the isotopic value in xylem water to those in soil pore water at different depths. However, such data are rare [@Evaristo:2016]. DNA barcoding of roots [@Jones.2011] may be used to estimate species-specific rooting depths or profiles, but DNA barcode libraries for tropical forests are still under development, and the method may be cost-prohibitive for the extent of sampling required. In general, community-scale data collection for rooting or water-sourcing depths is a formidable challenge in species-rich tropical forests.

\
  Characterizing the essential constraints to model species-rich communities [@Koven:2020; @Christoffersen:2016kd2d; @Wright:2010; @Bartlett.2019; @Lu.2020] entails identifying the topography of plant trait trade-offs in different environments, and how these relate to demographic rates (growth, recruitment and mortality). Significant efforts have been invested into identifying and linking universal drought indices, above-ground traits, and demographic rates. Such efforts have found correlations between vulnerability to embolism and hydraulic safety margins and mortality [@Anderegg:2016], although numerous counter examples also exist [@Hoffmann.2011; @Johnson:2018fad; @Nardini:2015; @Paddock:2013, @Venturas.2016]. Even though deep roots may mitigate hydraulic vulnerability and mortality-risk, hydraulic risk realized through hydrological (rather than meteorological) droughts--the outcome of the interaction between rooting depths, critical hydraulic thresholds and the actual vertical distribution of plant-available water--are not studied, especially in tropical regions.

\
  Here, we estimate plant-available water in the whole-soil column in a tropical forest, and inversely estimate rooting depths of co-occurring tree species from their growth responses, with a series of model calibrations and validation. We evaluate how rooting depth is linked to above-ground hydraulic traits and mortality rates over a 35 year period that experienced droughts of a variety of intensity, frequency and duration. Our hypotheses are that (i) deep-rooted trees have hydraulic traits associated with rapid water transport but cavitation-vulnerable xylem due to greater and more reliable water availability at depth, whereas shallow-rooted species are associated with less efficient but safer hydraulic traits due to more variable availability of shallow soil-water (see Table \@ref(tab:hypo)); and that, (ii) deep-rooted species have lower mortality rates during droughts due to a lower realized risk of hydraulic failure compared to shallow-rooted species. To our knowledge this is the first study to test for a mechanistic link between plant available water in the whole-soil column, trees' above- and below-ground hydraulic architecture and long-term mortality outcomes for a species-rich tropical forest.

# Materials and Methods

To obtain the daily dynamics of soil water potential in the whole-soil column, we locally parameterized and calibrated a 1-D hydrological water balance model. We then developed a novel, empirical inverse model to estimate species effective rooting depths (ERD) for large trees of `r length(erd.sp)` species (Fig. \@ref(fig:schematic)). We define ERD as the depth at which soil water potential best explained species-specific growth dynamics. The ERD model incorporated the impact of atmospheric and hydrological drought on growth, and was constrained with species-specific leaf vulnerability curves by using existing data for eight species, while developing trait-based proxies for the rest (based on data for a total of 21 species). We validated ERD against existing stable hydrogen isotope ratios ($\mathrm{\delta}$^2^H~xylem~) for a subset of six tree species as independent observations. We evaluated whether ERD was associated with above-ground hydraulic traits sourcing the latter from a set of rare datasets for six to seven species that overlapped with our assessment of ERD. We tested the relationship of ERD with xylem vulnerability to embolism, branch hydraulic safety margins, leaf turgor loss point and maximum stem-specific hydraulic conductivity. Finally, to test whether ERD explained risk of drought-induced mortality, we analysed species-specific large tree mortality dynamics over 35 years and estimated species-level realized hydraulic risk as the time a species’ spent beyond its critical hydraulic threshold.

## Study site description

We conducted this study at Barro Colorado Island (BCI), Panama. The entire island is forested and classified as a tropical moist forest in the Holdridge Life Zone system. Long-term hydrological monitoring at BCI began in 1972, while demographical monitoring began with the 50 ha ForestGEO plot establishment in 1981-82 [@Condit.1998uq; @Hubbell:1999; @Anderson-Teixeira:2014]. Rainfall at BCI is seasonal with a mean annual total of `r round(rain.man.stats$rain)` mm ($\pm$ `r round(rain.man.stats$rain.sd)` SD; `r min(clim$Year)`-`r max(clim$Year)`) and a pronounced dry season from mid-December through April with a mean total of 292 mm [@Paton.2020]. In the 50ha (1000 m × 500 m) old growth forest plot, all stems $\ge$ 1 cm diameter at breast height (dbh) are mapped, tagged with a unique number, identified to species and measured every five year through 1985-2015 for growth, mortality as well as recruitment of new stems into the 1 cm dbh size class [@Condit.2017]. This inventory represents 321 woody species, 28% of which are at least partly dry-season deciduous [@Condit:2000]. The elevation range in the plot is only 40 m (120-160 m) [@Harms:2001] and soils are homogeneous with red light clays accounting for 72% of the plot [@Baillie.2017]. The topsoil field texture is silty clay loam that gradually fines to silty clay in the subsoil. Soil is mostly free draining, but restricted subsoil permeability gives rise to temporary wet season ponding. Detailed descriptions of the climate, geology, flora and fauna of BCI can be found elsewhere [@Croat.1978; @Leigh.1982; @Gentry.1990].

## Model ELM-FATES for soil matric potentials

### Model description

To estimate dynamics of water availability by depth over the forest’s rooting zone, we used the Energy Exascale Earth System Land Model (ELM, @Caldwell.2019), coupled with the Functionally Assembled Terrestrial Ecosystem Simulator (FATES, @Koven:2020) (hereafter, ELM-FATES). ELM is a land-model that, among many features, simulates the physics and conservative dynamics of water, energy and carbon fluxes. In particular, soil hydrologic fluxes are vertically resolved among discrete soil layers (1-D) similar to that in CLM4.5 [@Bonan:2013]. FATES is a community-based, open-source model used for studying climate-vegetation interactions. FATES is a vegetation demography model, with a size-structured group of plants (cohorts) and successional trajectory-based patches based on the ecosystem demography [@Moorcroft.2001] approach. FATES couples to ELM by a common interface of water and carbon fluxes. Detailed descriptions for ELM and FATES can be found elsewhere [@Bisht.2018; @Fisher.2010; @Fisher.2015; @Koven:2020].

\
  ELM-FATES simulates interception, throughfall, canopy drip, infiltration, evaporation, surface runoff, subsurface drainage, redistribution within the soil column, and groundwater discharge and recharge so as to simulate changes in canopy water $\mathrm{\Delta}W_{\textrm{can}}$, surface water $\mathrm{\Delta}W_{\textrm{sfc}}$, soil water $\mathrm{\Delta}W_{\textrm{liq},z_{i}}$ at depth $z_{i}$ and water in an unconfined aquifer $\mathrm{\Delta} W_{a}$ (omitting processes relevant to snow, wetlands or lakes). Conservation of these terms is expressed as follows:

\begin{equation}
\begin{split}
\mathrm{\Delta}W_\textrm{can} + \mathrm{\Delta}W_\textrm{sfc} + \sum_{i=1}^{15} \mathrm{\Delta}W_{\textrm{liq},z_{i}} + \mathrm{\Delta}W_{a} = \left(q_\textrm{rain} - E_{v} - E_{g} - q_\textrm{over} - q_\textrm{h2osfc} - q_\textrm{drai}\right)\mathrm{\Delta}t
\end{split}
(\#eq:water-balance)
\end{equation}
where, $q_{\textrm{rain}}$ is rainfall, $E_{v}$ is transpiration of the forest, $E_{g}$ is ground evaporation, $q_{\textrm{over}}$ is surface runoff, $q_{\textrm{h2osfc}}$ is runoff from surface water storage, $q$~drai~ is subsurface drainage (all in mm H~2~O) and $\mathrm{\Delta}t$ is the time-step (here, 1 hr). The soil profile is discretized into maximum 15 exponentially distributed soil layers with layer node depth $z$. Here, $z$ $\in$ $Z$; $Z$ = (0.01, 0.03, 0.06, 0.12, 0.21, 0.37, 0.62, 1, 1.7, 2.9, 4.7, 7.8, 13) m.

### Model parameterization

To parameterize ELM-FATES, we identified eleven parameters relevant for Eq. \@ref(eq:water-balance), determined their ranges based on literature for the study site, else for the tropics (Table \@ref(tab:param)), and ran 5000 simulations using Latin Hypercube Sampling [LHS; @Stein.1987] from this global parameter space. In the Conrad catchment soil hydraulic conductivity decreases with depth from 12.5 and 60 cm (@Godsey:2004; Table \@ref(tab:param)). We generated 5000 LHS samples within the 95% CI for the two depths and assumed a linear decline between each paired sample, while for depths <12.5 cm and >60 cm, we assumed the same value as that for depths 12.5 and 60 cm, respectively (Fig. S\@ref(fig:ksat-profiles)). For greater accuracy, instead of the ELM default pedo-transfer functions, we estimated parameters of soil retention curves using existing data for gravimetric water content (GWC) vs. $\mathrm{\Psi}$~soil~ for depths of 0.15, 0.4 and 1 m [@Kupers:2019]. We converted GWC to volumetric (VWC, $\theta$) using a measured bulk density value of 0.8 g cm^-3^ (see below). A @Campbell.1974 empirical equation was fitted for all depths pooled together, as exploratory analyses indicated differences in water retention among depths were nominal,

\begin{equation}
\mathrm{\theta}_{\textrm{soil}} = \mathrm{\Psi}_e\left(\frac{\mathrm{\theta}}{\mathrm{\theta}_{s}}\right)^{-b}
(\#eq:soil-ret-curve)
\end{equation}
where $\mathrm{\Psi}_e$ is the air-entry matric potential, $\mathrm{\theta}_{s}$ is the saturated volumetric water content, and $b$, an index for soil pore-size distribution, respectively. $\mathrm{\Psi}_e$, $\mathrm{\theta}_{s}$ and $b$ that best-fit the data were 200 (mm H~2~O), 0.51 (cm^3^ cm^-3^) and 10 (unitless), respectively.

\
We ran ELM-FATES for the 5000-member parameter ensemble with hourly climate drivers measured at the BCI meteorological station over 1985-2018 initialized with the observed stand structure from the 50 ha plot [@Condit.2019a] in a single site mode (see https://biogeodb.stri.si.edu/physical_monitoring/research/barrocolorado for meteorological data and @Faybishenko.2018 for QA/QC procedures). As our key interest here was on deriving soil water availability, we ran ELM-FATES in a lower-complexity configuration: static stand structure, and with a single plant functional type (PFT) of evergreen trees. The latter was chosen as only 9.7% of BCI crown area is dry-season deciduous [@Condit:2000] and addition of a dry-deciduous PFT did not significantly alter results (not shown). Holding the stand structure static rather than allowing it to emerge from physiologically-mediated competition between trees removes a set of internal model feedbacks and thus the only differences between model simulations are driven by direct physiological effects of perturbed model parameters.

### Model calibration

We calibrated ELM-FATES over 2012-2018 against three key fluxes and states in Eq. \@ref(eq:water-balance), namely, (1) evapotranspiration ET from the eddy flux tower by the 50 ha plot ($\backsimeq$ $E_{v} + E_{g}$ in Eq. \@ref(eq:water-balance); 2012-2017; see [SI](#SIData.matteo))), (2) local stream discharge $q_\textrm{run}$ ($\backsimeq$ $q_\textrm{over} + q_\textrm{drai}$; 2012-2018; see @Paton.2019 and [SI](#SIData.steveP)) and (3) soil volumetric water content (VWC) $\mathrm{\theta}_{z}$ ($\backsimeq$ $W_{\textrm{liq,} z}$) from two sources: (i) a long-term (2012-2018) record of VWC $\mathrm{\theta}_\textrm{0-0.15 m}$ averaged across three vertical time domain reflectometry (TDR) probes over depth 0-15 cm from three locations near the eddy flux tower, and (ii) plot-wide snap-shot measurements of VWC during the dry season of 2015 and 2016 at depths of 0.15, 0.4 and 1 m (1299 samples covering all soil types and habitats; see [SI](#SIData.VWC) and @Kupers:2019 for details; hereafter, $\mathrm{\theta}_{\textrm{0-0.15 m}}$, $\mathrm{\theta}_{\textrm{0.4 m}}$ and $\mathrm{\theta}_{\textrm{1 m}}$ refer to plot-wide averages).

\
 For ELM-FATES calibration we calculated an objective function for each of the 5000-member ensemble by equally weighting standardized RMSE between observations and simulations across all fluxes and states. We describe the objective function $O$ as :

\begin{equation}
\begin{split}
\overline{O} = \frac{1}{6}\sum \textrm{RMSE}^{*}\textrm{(ET,}E_{v} + E_{g}) + \textrm{RMSE}^{*}(q_{\textrm{run}},q_{\textrm{over}} + q_{\textrm{drai}}) + \textrm{RMSE}^{*}(\mathrm{\theta}_{\textrm{0-0.15~m}}, {W_{\textrm{liq,}\overline{z}}}) \\
& + \textrm{RMSE}^{*}(\mathrm{\theta}_{\textrm{0.15~m}}, W_{\textrm{liq,`r soil.depths[4]`m}}) + \textrm{RMSE}^{*}(\mathrm{\theta}_{\textrm{0.4~m}}, W_{\textrm{liq,`r soil.depths[6]`~m}}) + \textrm{RMSE}^{*}(\mathrm{\theta}_{1~m}, W_{\textrm{liq,`r soil.depths[8]`~m}})
(\#eq:obj-func)
\end{split}
\end{equation}

based on monthly sums for both ET and $q$~run~ and daily averages for $\mathrm{\theta}$s. $W_{\textrm{liq,}\overline{z}}$ is the average of daily $W_{\textrm{liq,} z}$, where $z$ $\in$ $Z$; $Z$ = (`r paste(soil.depths[1:4], collapse=", ")`) m. Superscript * indicates that RMSE is standardized between 0 and 1 across the 5000 simulations.

\
Preliminary analyses showed that simultaneously fitting the model to all states and fluxes did not capture the soil moisture dynamics at 1 m depth well, our key priority for model calibration. To ensure model fit at 1 m depth, we first chose 300 parameter ensemble members that minimized $\textrm{RMSE}^{*}({\mathrm{\theta}_{\textrm{1 m}}}, W_{\textrm{liq,`r soil.depths[8]` m}})$ (< 0.2), and among those 300 ensemble members we chose 100 with the lowest values for the objective function $\overline{O}$. See Table \@ref(tab:param) for the ranges of best-fit values for different parameters.

### Model evaluation

We evaluated ELM-FATES by calculating RMSE between simulated and observed long-term daily VWC for the depths of 0.1, 0.4 and 1 m (2016-2018), based on a dataset we had left-out during calibration. We obtained these observations from three horizontal TDR probes at the depths of 0.1, 0.4 and 1 m at a location near the vertical probes (see [SI](#SIData.VWC)).

### Hydrological droughts

We ran ELM-FATES with the chosen 100 ensemble members from 1985-2018, with the first 5 years used for model spin-up. Extreme hydrological droughts were identified by depth $z$ as days for which $\mathrm{\Psi}_{\textrm{soil},z}$ was more negative than the 5^th^ percentile of $\mathrm{\Psi}_{\textrm{soil},z}$ for a given day of the year.

## Model for effective rooting depth (ERD)

### Model description

To inversely estimate effective rooting depth, we tested alternate empirical models for species-specific 5-yearly census growth, which use effective rooting depth as an input (Eq. \@ref(eq:M-chosen) and those in SI). Many factors limit plant growth, including soil moisture, VPD, leaf area seasonality and radiation [@Lawrence.2019]. Because pre-dawn $\mathrm{\Psi}$~leaf~ generally approaches $\mathrm{\Psi}$~soil~, we used $\mathrm{\Psi}$~soil~ in leaf hydraulic vulnerability curves to predict maximum diurnal $K$~leaf~, which we assumed controls photosynthetic rate and growth (@Brodribb.2002; @Brodribb.2017; see Table \@ref(tab:symbols) for trait symbols and their definitions). We assumed $K_\textrm{leaf}$ to be a hydraulic bottleneck in the plant, and thus to determine or co-determine whole plant hydraulic response to soil drying; notably, root hydraulic vulnerability may also be a strong driver, and recent studies suggest that its vulnerability would be similar to, or correlated with that of leaves (see @Cochard.2020 and references therein). $K_\textrm{leaf}$ in our models, being based on vulnerability curves, is dynamic and thus depends on rooting depth and rhizosphere water potential. The use of maximum diurnal $\mathrm{\Psi}$~leaf~ based on pre-dawn $\mathrm{\Psi}$~soil~ as a proxy for diurnal $\mathrm{\Psi}$~leaf~ and its influence on growth at diurnal time-scales (to relate to 5-year average growth values) thus does not consider diurnal dynamics of $\mathrm{\Psi}$~leaf~ and other factors that would play a role in such diurnal dynamics, such as the buffering effect of stem water storage capacitance on $\mathrm{\Psi}$~leaf~, and may thus overestimate ERD (see Discussion). VPD affects growth non-linearly, with growth increasing with VPD up-to a VPD threshold, then decreasing as leaf pores (stomata) close reducing water-uptake [@Grossiord.2020; @Yang.2019]. To account for the non-linear impact of VPD on growth, we used predicted gross primary productivity (GPP; hereafter, $\hat{\textrm{VPD}}$) from a locally derived polynomial relationship between GPP and VPD (see SI). Apart from stomatal control, leaf-drop may further limit water-uptake and growth. We therefore regressed species-specific growth against multiplicative or additive effects of one or more factors calculated daily and averaged over 5-yearly censuses: $\hat{\textrm{VPD}}$, $K_\textrm{leaf}$ and the leaf area index (LAI), all standardized between 0 and 1. Incorporation of radiation did not improve model fitting, possibly due to the coincidence of higher temperature, higher radiation, and lower humidity during the dry season at BCI.

\
 For the best model (see model validation and selection below), daily average growth for species $s$, $G_{s,t|z}$, in the census interval $t$ for depth $z$ and for a given hydrological realization $h$ is described as follows:

\begin{equation}
G_{s,t|z} = \beta_{0,s|z,h} + \beta_{1,s|z,h}(\frac{1}{n_{t}}\sum_{i=1}^{n_{t}} K_{\textrm{leaf},s,i|z,h}^{*} \hat{\textrm{VPD}}_{i}^{*}) + \epsilon_{s,t|z,h}
(\#eq:M-chosen)
\end{equation}
where n~t~ is the total number of days in census interval $t$, the superscript * indicates that the variable has been standardized between 0 and 1 (within species for $K_{\textrm{leaf}}$), $\beta_{0}$ and $\beta_{1}$ are model coefficients and $\epsilon$ is the model error term. See SI for all the alternate models tested.

\
 For each species, we estimated diurnal maximum value of $K_{\textrm{leaf},s,i|z,h}$ for a species $s$ by substituting $\mathrm{\Psi}_{\textrm{soil},s,i|z,h}$ in the leaf hydraulic vulnerability curves defined below in Eq. \@ref(eq:vc-exp) using species-specific parameters $A_{s}$ and $B_{s}$ as:

\begin{equation}
K_{\textrm{leaf},s,i|z,h} = A_{s}e^{-B_{s}{\mathrm{\Psi}}_{\textrm{soil},i|z,h}}
(\#eq:k-soil)
\end{equation}
 Checking through each $z$, we estimated species effective rooting depths (ERD), defined as the depth $z$ at which soil water dynamics ($\mathrm{\Psi}_{\textrm{soil,} z}$) best explained observed dynamics of species growth. Using Eq. \@ref(eq:M-chosen), for a species $s$ and hydrological realization *h*, we obtained Akaike's Information Criterion (AIC) and Pearson's correlation coefficient $r$ for each soil depth $z$. For each *h*, we retained depths with the maximum likelihood (similar AIC based support, see below), and amongst them selected the best-fit depth as the one with the maximum Pearson's $r$, if $r$ >= 0.71. For a species $s$, we defined effective rooting depth (ERD) as the median of the best-fit depth over all *h*, which can be described as :

\begin{equation}
\textrm{ERD}_{s} = \underset{\forall \, h \, \in \, H}{\textrm{med}} \Big(z\,\Big|\, \Big(r_{s|z,h} = \underset{\forall \, z \,\in\,Z}{\textrm{max}}(r_{s|z,h}| \mathrm{\Delta}_{s|z,h} \le 2\Big)\Big) \, \bigcap \, r_{s|z,h} \ge 0.71\Big)
(\#eq:erd)
\end{equation}

where $\mathrm{\Delta}_{s|z,h} = \textrm{AIC}_{s|z,h} - \underset{\forall \, z \,\in\,Z}{\textrm{min}}(\textrm{AIC}_{s|z,h})$,  $\textrm{AIC}_{s|z,h} = -2\textrm{log}(\mathcal{L}(\hat{\beta}_{s|z,h}|{G_{s}})) - 2k$, $\mathcal{L}$ is the likelihood function, $G$ the observed growth time series with $k$ observations (or, census intervals $t$). $H$ = (1, .., 100), the 100 best-fit hydrological realizations and $Z$ = (0.4, `r paste(soil.depths[8:length(soil.depths)], collapse=", ")`) m. We removed finely resolved soil layers <0.21 m and averaged $\mathrm{\Psi}_{\textrm{soil},z}$ for depths 0.21, 0.37 and 0.62 (=0.4) m as sufficient resolution for model validation.

#### Growth data

For diameter growth estimates, to minimize the effect of light variation among trees, we selected only large trees ($\ge$ 30 cm dbh) in the 50 ha plot and also species whose maximum height was $\ge$ 30 m (hereafter, canopy species) and thus are likely to be fully exposed to the sun. Calculating individual tree growth rates across six 5-yearly censuses (1990-2015; @Condit.2019a; @Condit.2019b), removing outliers, obtaining residuals from a dbh model of growth to account for the size effect on growth (see SI), we estimated species $s$ growth time series $G_{s}$ (cm yr^-1^) as the median of standardized dbh model residuals, for only those species ($n$ = `r length(erd.sp)`) with complete records for at least three trees (median `r median(g.n$n)`, max `r max(g.n$n)` trees per species).

#### Leaf hydraulic vulnerability curves

We obtained leaf hydraulic vulnerability curves ($K_\textrm{leaf}$ vs. $\mathrm{\Psi}$~leaf~) for adult trees of 21 common species from Zailaa *et* *al.* (to be submitted; see SI for brief description of methods) described as:

\begin{equation}
K_{\textrm{leaf},s} = A_{s}e^{-B_{s}{\mathrm{\Psi}}_{\textrm{leaf}_{s}}} + \epsilon_{s}
(\#eq:vc-exp)
\end{equation}
where $A$ and $B$ are fitted species-specific parameters, $\epsilon_{s}$ an error term. These 21 species included eight of the `r length(erd.sp)` species selected for ERD estimation. For the rest `r sp.n.model` species, we obtained vulnerability curves using trait-based proxies. We identified scaling relationships between fitted parameters $A$ and $B$ in Eq. \@ref(eq:vc-exp) and two traits; namely, WSG, the wood specific gravity, and LMA, the leaf mass per unit area (@Wright:2010; see SI for details). We fitted polynomial equations described as,

\begin{equation}
\begin{split}
B_{s} = `r bcf[1]` `r bcf[2]`\textrm{WSG}_{s} + `r bcf[3]`\textrm{WSG}_{s}^2 `r bcf[4]`\textrm{WSG}_{s}\textrm{LMA}_{s} \\
& + `r bcf[5]`\textrm{LMA}_{s} `r bcf[6]`\textrm{LMA}_{s}^2 + \epsilon_{s,b}
(\#eq:b-mod)
\end{split}
\end{equation}

\begin{equation}
A_{s} = `r acf[1]` `r acf[2]`B_{s} `r acf[3]`B_{s}^2 + `r acf[4]`B_{s}\textrm{LMA}_{s} + `r acf[5]`\textrm{LMA}_{s} `r acf.6`\textrm{LMA}_{s}^2 + \epsilon_{s,a}
(\#eq:a-mod)
\end{equation}
where $\epsilon_{s,a}$ and $\epsilon_{s,b}$ are error terms.

\
As these fits explained a large proportion of variation in parameters $A$ and $B$ (see results), we sequentially used Eq. \@ref(eq:b-mod) and \@ref(eq:a-mod) to predict parameters $B$ and $A$, respectively, for the `r sp.n.model` ERD species without direct data and estimated leaf hydraulic vulnerability curves using Eq. \@ref(eq:vc-exp). We thus obtained parameters $A$ and $B$ for all the `r length(erd.sp)` ERD species for use in Eq. \@ref(eq:b-mod). We also estimated species $\mathrm{\Psi}$~50,leaf~ from their vulnerability curves.

#### Leaf Area Index (LAI)

In some of the alternative models for effective rooting depth, we explored the effect of seasonality in leaf area index (LAI) on growth. We assumed species-specific mean seasonal curves for LAI (standardised between 0 and 1; unitless), informed by a combination of long-term records for leaf-fall [@Wright.1990aeo] and leaf-lifetime (@Osnas.2018; see SI).

### Model validation and selection

We used $\mathrm{\delta}$^2^H~xylem~ as an independent observation to validate the ERD models. As root water-uptake is generally a non-fractionating process, tree $\mathrm{\delta}$^2^H~xylem~ reflects the signature of source water. Given a vertical gradient of $\mathrm{\delta}$^2^H in soil and groundwater $\mathrm{\delta}$^2^H~xylem~ provides an index of rooting depth [@Dawson.1991]. We leveraged $\mathrm{\delta}$^2^H~xylem~ from BCI [@Meinzer:2001] for the dry-season of March 1997 as this period showed largest seasonal divergence in $\mathrm{\delta}$^2^H~xylem~ among species and vertically in soil and groundwater $\mathrm{\delta}$^2^H at natural abundance level (Fig. S\@ref(fig:iso-data)). Analyzing $\mathrm{\delta}$^2^H~xylem~ for 12 tree species @Meinzer:1999 found that four species sourced water deeper than 1 m, two other at least 0.8 m deep, one between 0 to 0.2 m, and five in the 0.2 to 0.7 m depth range over which the relationship between soil depth and $\mathrm{\delta}$^2^H~soil~ was variable.

\
We removed those species that would be leafless in March 1997 because $\mathrm{\delta}$^2^H~xylem~ for these species may not be linked with water sourced at the time of measurement and thus not comparable to species that had leaves. We also removed one species whose $\mathrm{\delta}$^2^H~xylem~ was in the uncertain depth range. For each model of ERD (Eq. \@ref(eq:M-chosen) and alternatives in SI), modeled species ERD was regressed against species $\mathrm{\delta}$^2^H~xylem~ for a maximum of six species.

## Relationships between ERD and above-ground hydraulic traits

We evaluated whether ERD was associated with above-ground hydraulic traits sourcing the latter from a set of rare datasets [@Wolfe.2019; @Wolfe.2020] for six to seven species that overlapped with our assessment of ERD. We regressed species ERD against maximum stem hydraulic conductivity $K_\textrm{max,stem}$ (*n*=7), leaf turgor loss point $\mathrm{\Psi}$~tlp~ (*n*=7), vulnerability to embolism from cavitation $\mathrm{\Psi}$~88,stem~ measured in terms of pressure at which 88% of $K_\textrm{max,stem}$ is lost (*n*=7), and hydraulic safety margins $\mathrm{\Psi}$~min~ - $\mathrm{\Psi}$~88,stem~ (*n*=6). See [SI](#SIData.brett) and Table S\@ref(tab:hyd-table) for data collection and estimation of these variables.

## Mortality analyses and species-specific drought experience

To test whether ERD plays a role in mitigating mortality risk, we calculated mortality rates for large trees (here, $\ge$ 10 cm dbh) in the 50ha plot for species with ERD estimates (all canopy species with maximum height $\ge$ 30 m) as well as average abundance of at least 20 trees in the plot ($n$ = `r length(erd.mrate.sp)`) [@Condit.2019a]. For each of the seven census intervals $t$ in the 35-yr record (1981, 1985, 1990, 1995, 2000, 2005, 2010, 2015), mortality rate, $\textrm{M}_{t}$ (% yr^-1^) for species $s$ for two successive censuses, $c_{1}$ and $c_{2}$, was calculated as $\textrm{M}_{s,t} = \frac{D_{s,c_{2}}}{N_{s,c_{1}}}\times\frac{100}{\textrm{d}}$, where $N_{s,c_{1}}$ and $D_{s,c_{2}}$ are the total number of large trees of species $s$ present in the 50 ha plot in $c_{1}$ and dead in $c_{2}$, respectively, and $d$ is the duration based on mean dates of $c_{1}$ and $c_{2}$.

\
We regressed species ERD against species $\textrm{M}_{s,t}$ for each census interval $t$ for the whole pool of ($n$ = `r nrow(mrate.depth.mean)`) species with $\textrm{M}_{s,t}$ estimates and for evergreen species alone ($n$ = `r nrow(mrate.depth.mean %>% subset(deciduous == "E"))`). To assess whether deeper ERD species escaped the experience of significant hydrological drought, we calculated realized hydrulic risk for each species as the proportion of days in each census interval $t$ of the $\mathrm{\Psi}_\textrm{soil}$ simulation period (1990-2015) during which soil water potential in the soil layer matching species ERD, $\mathrm{\Psi}_{\textrm{soil},z=\textrm{ERD}}$, were more negative than species $\mathrm{\Psi}$~crit~. $\mathrm{\Psi}$~crit~ was defined as $\mathrm{\Psi}$~50,leaf~.

\
All statistical analyses were conducted in the *R* statistical environment (version 3.5.3, @R:2019).

# Results

## Soil water dynamics by depth and hydrological droughts

The 100 ELM-FATES ensemble members with the best fits to observed soil moisture dynamics captured soil moisture seasonality at multiple depths (Fig. \@ref(fig:hydro-calib)a, b), including the out-of-sample observations. These simulations also captured the dynamics in stream discharge (Fig. \@ref(fig:hydro-calib)c) and in evapotranspiration (Fig. \@ref(fig:hydro-calib)d), slightly underestimating peak discharge in wet years and slightly overestimating peak ET. The reduction in parameter range (Table S\@ref(tab:param)) in the best-fit ensembles compared to the tested global ranges showed that the model calibration was sensitive to the Ball-Berry stomatal slope parameter, the ELM root distribution parameter that regulates the depth of the rooting profile, soil hydraulic conductivity profile especially at depth, and the adjustment factor that modifies soil hydraulic conductivity to account for macroporosity and direct flow paths (Fig. S\@ref(fig:sens-sm-et), S\@ref(fig:sens-run)).

\
Soil water potential dynamics $\mathrm{\Psi}_{\textrm{soil},z}$ by depth z from 0.01 through 13 m obtained from the best-fit simulations revealed that every 5-yr census interval had an extreme hydrological drought year, but census intervals varied in terms of number of extreme years, drought intensity at depth, drought seasonality and duration (Fig. \@ref(fig:swp-droughts)). Hydrological droughts in census intervals 2000-2005 and 2005-2010 distinctly occurred in the wet season, with one to two extreme hydrological drought years each, depending on the depth location (Fig. S\@ref(fig:swp-droughts)). Across the simulation period 1990-2015, simulated $\mathrm{\Psi}_{\textrm{soil},z}$ remained above -0.5 MPa from depths 1.7 to 13 m (Fig. S\@ref(fig:swp-all-depths)).

## Leaf hydraulic vulnerability curves

Parameters of leaf vulnerability curves were predictable from WSG and LMA (Fig. \@ref(fig:ab-fit)). WSG and LMA explained a large proportion of variance in parameter $B$ (Eq. \@ref(eq:b-mod); Adj. $R$^2^ = `r bfit.r2`; $p$ `r  bfit.p`). WSG and parameter $B$ explained a large proportion of variance in parameter $A$ (Eq. \@ref(eq:a-mod); *Adj. R*^2^ = `r afit.r2`; $p$ `r  afit.p`; also see Fig. S\@ref(fig:obs-plc-curves)). This predictive power allowed us to estimate leaf vulnerability curves used in the ERD models for `r sp.n.model` out of `r length(erd.sp)` ERD species that lacked direct observations (Table S\@ref(tab:ab-param), Fig. S\@ref(fig:obs-comm-plc-curves)).

## Effective rooting depths (ERD)

The best ERD model (with a multiplicative effect of VPD and not LAI, Eq. \@ref(eq:M-chosen)) explained a large fraction of the variance in $\mathrm{\delta}$^2^H~xylem~ ($R$^2^ = `r lm.erd.iso.r2`; $p$ = `r lm.erd.iso.p`, $n$ = `r length(unique(erd.iso$sp))`; Fig. \@ref(fig:erd-iso)). Two of the alternative models with LAI had a strong relationship with $\mathrm{\delta}$^2^H~xylem~ as well, but one that was based on a smaller set of species ($n$ = 4), because fewer species passed the criteria of sufficient species-level goodness-of-fit for these models (<20 species vs. `r length(erd.sp)` species for the best model; Fig. S\@ref(fig:other-erd-mods)). For the rest of the analyses we therefore used ERD estimates from the less parsimonious, best model.

\
Modeled effective rooting depths (ERD) for the `r length(erd.sp)` large, canopy tree species varied from `r min(df.erd.to.plot$depth)` m to `r max(df.erd.to.plot$depth)` m (Fig. \@ref(fig:sp-erd)). Notably, two species, *Luehea seemannii* and *Trichilia tuberculata*, that @Meinzer:1999 found to have $\mathrm{\delta}$^2^H~xylem~ values between soil water and groundwater, suggesting that these species sourced most of the water from depths >1 m, were also identified by our model with ERD >1 m (2.9 m for both species; Fig. \@ref(fig:erd-iso)).

## Relationships between ERD and above-ground hydraulic traits

Species with deeper ERD showed greater $K_\textrm{max,stem}$ (Spearman's' $r$ = `r cor.k`, $p$ = `r p.k`; Fig. \@ref(fig:erd-hyd)a), less negative leaf $\mathrm{\Psi}$~tlp~ ($r$ = `r cor.tlp`, $p$ = `r p.tlp`; Fig. \@ref(fig:erd-hyd)b), less negative $\mathrm{\Psi}$~88,stem~ and thus greater vulnerability to embolism from cavitation ($r$ = `r cor.stem.88`, $p$ = `r p.stem.88`; Fig. \@ref(fig:erd-hyd)c), and narrower hydraulic safety margins ($\mathrm{\Psi}$~min~ - $\mathrm{\Psi}$~88,stem~; $r$ = `r cor.hsm`, $p$ = `r p.hsm`; Fig. \@ref(fig:erd-hyd)d).

## Effective rooting depths, mortality and hydrological droughts

Among evergreen species, ERD explained `r m.evg.r2.90`, `r m.evg.r2.05` and `r m.evg.r2.10`% variation in mortality rates during census intervals 1985-1990, 2000-2005 and 2005-2010, respectively, such that species mortality rates significantly decreased with deeper ERD ($p$ = `r m.evg.p.90`, `r m.evg.p.05` and `r m.evg.p.10`, respectively; Fig. \@ref(fig:erd-mort-census-evg)). When deciduous species were included, ERD explained lower but still significant proportion of variation in mortality in those intervals ($R$^2^ = `r m.r2[1]`, `r m.r2[2]` and `r m.r2[3]`% and $p$ = `r m.p.90`, `r m.p.05` and `r m.p.10`, respectively; Fig. \@ref(fig:erd-mort-census)). Census intervals 2000-2005 and 2005-2010 were also the periods during which extreme soil droughts distinctly occurred in the wet season, effectively extending the dry season into the wet season (Fig. \@ref(fig:swp-droughts)). While our $\mathrm{\Psi}_{\textrm{soil}}$ estimates did not extend prior to 1990, due to a lack of reliable climate forcing, a prior study found that during the period 1985-1990, for which ERD explained significant mortality, BCI experienced three extreme deficit water years, following the 1983 El-Niño year [@Condit.2017]. Analysis of a species' realized hydraulic risk as the time spent under significant hydrological drought (% days $\mathrm{\Psi}_{\textrm{soil},z=\textrm{ERD}}$ < $\mathrm{\Psi}$~crit~) showed that species with shallower ERD experienced greater realized hydraulic risk and increasingly so over 2000-2005 and 2005-2010 (Fig. \@ref(fig:erd-time)). For the census interval 1990-1995, species with the shallowest ERD also showed greater realized hydraulic risk (Fig. \@ref(fig:erd-time)), but during the dry season (; Fig. \@ref(fig:swp-droughts), and without elevated mortality rates (Fig. \@ref(fig:erd-mort-census). Overall, species with ERD deeper than 1.7 m did not cross species $\mathrm{\Psi}$~crit~ for the entire 25-yr period for which $\mathrm{\Psi}_{\textrm{soil},z}$ was modeled (1990-2015). Soil depths deeper than 1 m remained above `r round(psi.1.7.min$median, 2)` MPa (`r round(as.numeric(psi.1.7.min$q97.5), 2)`, `r round(as.numeric(psi.1.7.min$q2.5), 2)`; median, 95% CI) through all the dry seasons and drought periods, which is above $\mathrm{\Psi}$~crit~ of the most sensitive tree species (-0.28 MPa).

# Discussion

Our novel approach for estimating ERD using 25 years of tree growth, species-specific leaf vulnerability curves, modeled soil moisture profiles, and observed VPD was consistent with estimates using $\mathrm{\delta}$^2^H~xylem~. Our analyses revealed that co-occurring large canopy tree species with deeper ERD were associated with higher stem hydraulic efficiency, but not safety (Fig. \@ref(fig:erd-hyd)). Species with deeper ERD showed higher maximum stem hydraulic conductance, and greater xylem vulnerability to embolism and narrower hydraulic safety margins compared to shallow ERD species. Nevertheless, deeper ERD species showed lower mortality rates than shallower ERD species in significant wet season droughts over the 35-yr record (1981-2015) (Fig. S\@ref(fig:erd-mort-census)), especially amongst the evergreen species (Fig. \@ref(fig:erd-mort-census-evg)). Analysis of time spent beyond $\mathrm{\Psi}$~crit~ in the soil zone matching ERD $\mathrm{\Psi}_{\textrm{soil},z=\textrm{ERD}}$ showed that this was likely due to decreasing realized risk of hydraulic failure with increasing access to water for species with deeper ERD (Fig. \@ref(fig:erd-time)). We thus demonstrate, for the first time, a link between realized drought risk, hydraulic architecture and long-term mortality outcomes for large tree species of a tropical forest.

## Realized drought risk integral to assessing drought-vulnerability

Our study brings attention to the need for assessing drought sensitivity in terms of species realized drought exposure by accounting for hydrological drought and trees' rooting depths. We found that species accessing deeper water had greater xylem vulnerability to embolism and narrower branch hydraulic safety margins (Fig. \@ref(fig:erd-hyd)), that is, traits that are commonly identified as proxies for mortality risk [@Anderegg:2016], but they were less droughted and had lower mortality compared to species relying on shallower water through periods of extreme droughts (Fig. \@ref(fig:erd-mort-census-evg), S\@ref(fig:erd-mort-census), \@ref(fig:swp-droughts)). Deeper ERD species on average had more hydraulically efficient, but risky, suite of above-ground traits optimized for an environment of more reliable and greater water source at depth that resulted in a lower mortality risk. By contrast shallow ERD species had inefficient, but safer, xylem likely optimized for high variability in $\mathrm{\Psi}_\textrm{soil}$ in the shallow soil layers, and resulting in higher variance in mortality rates across censuses as $\mathrm{\Psi}_{\textrm{soil},z\textrm{ = ERD}}$ crossed ERD $\mathrm{\Psi}_\textrm{crit}$ (Fig. \@ref(fig:erd-mort-census-evg); Fig. \@ref(fig:erd-time)). Our results are consistent with studies that included three or four species each with known rooting or water-sourcing depths and assessed their hydraulic traits and mortality rates during extreme droughts [@Johnson:2018fad; @Nardini:2015; @Venturas.2016; also see @Brum.2019; @Brum.2017]. Globally, large trees tend to exhibit greater growth reductions, lower post-drought resilience, and greater increases in mortality relative to their understory counterparts [@Bennett:2015; @Phillips:2010]. Our finding that deep-water access buffers drought-induced mortality in large trees is relevant for understanding drought resistance, resilience and recovery of this important functional group [@Bennett:2015; @McGregor.2020]

\
Our result--of lower mortality in deep-rooted trees--contrasts with the inverse model finding of @Chitra-Tarak.2018 (hereafter, CT2018) in which deeper ERD species in a South-Asian seasonally tropical dry forest had higher mortality in a rare, prolonged drought. The hydrological model in CT2018 revealed that the multi-year drought exhausted the deep soil and even bedrock water availability (also see [@Goulden.2019; @Ivanov:2012]). In contrast, our hydrological modeling at BCI finds that $\mathrm{\Psi}$~soil~ for depths deeper than 1 m did not cross $\mathrm{\Psi}$~crit~ of the most sensitive tree species (-0.28 MPa) for all the dry seasons and droughts during 1990-2015. Mean annual rainfall of 1095 mm in CT2018 compared to `r round(rain.man.stats$rain)` mm at BCI, and precipitation to potential evapotranspiration ratio of nearly one in CT2018 compared to nearly two in BCI are major factors in the different mortality responses. At BCI deep soil layers were recharged annually (Fig. S\@ref(fig:swp-all-depths), \@ref(fig:swp-droughts)), whereas in CT2018 they were not. The contrast between the two studies highlights the combined role of seasonal precipitation input and site-specific hydrology in modulating mortality-risk of deep-rooted species.

## Effective rooting depths at the tree community level

By estimating ERD for large trees of `r length(erd.sp)` canopy tree species, we make an important advance in modeling effective rooting depths at the community level in species-rich tropical forests. Our chosen ERD model is a key improvement over the model of CT2018 ($n$ = 12 species) as we employ a physiologically meaningful representation, fewer parameters and corroboration with tree and soil data for $\mathrm{\delta}$^2^H. We note that stable water isotopes for plant and below-ground sources are critical for calibration or validation of ERD models for scaling upto community, although the interpretation of isotopic data is still under research [@Deurwaerder.2020; @Adams.2020; @Bowers.2020]. As in CT2018, we also interpret our ERD estimates as revealing relative differences among species' effective rooting depths rather than absolute depths. As our ERD model ignores stem water storage capacitance [@Wolfe.2017] it may have overestimated ERD especially in those species for which capacitance is important as a drought-avoidance strategy, for example, the deciduous species [@Borchert.2005jma].

## Future directions

The relationships that we identify between above- and below-ground traits, the water environment vertically below-ground and mortality rates are important for representing diversity in DGVMs, which intrinsically rely on the parameterization of contrasting life history strategies [@Scheiter:2013] and the simulation of competition between those strategies. We found that LMA and WSG were strong predictors of leaf hydraulic vulnerability curves. This offers the promise of a greater ability to parameterize the 'hard' hydraulic traits with the abundant 'soft' trait data, thus allowing for a better representation of forest hydrodynamics. The relationships between ERD and above-ground hydraulic traits that we find will provide an important insight on how to model rooting depths and their coordination or trade-offs with other traits in order to better represent the functional diversity of tropical forests, and their trajectories into the future.

\
Our inverse ERD model was parameterized on 5-year growth data, with five data points over a 25-yr period, which decoupled climate events and demographic outcomes. Future studies could better constrain the ERD model with higher frequency growth data such as those from dendrometer bands. At high temporal resolution, however, the role of reversible dehydration in tree diameter change increases [@Chitra-Tarak.2016; @Chitra-Tarak.2015; @Mencuccini.2017], but that may provide an avenue to include stem water storage  and dynamic rooting depths in ERD models. Three of our exploratory ERD models included LAI seasonality, but we did not select them as they worsened the fit with growth data for many species (Fig. S\@ref(fig:other-erd-mods)). We interpret this result as that VPD and leaf hydraulic vulnerability curves may be adequate to explain inter-census differences in growth (via stomatal control), but also acknowledge that our estimates of seasonality of LAI is a tentative estimate that combines leaf-fall data and the timing of leaf-gain backtracked from leaf lifetime, and omits inter-annual variation. Future studies may improve models of LAI seasonality and make use of data from new technologies such as drone based monitoring of LAI in species-rich tropical forests [@Park.2019].

\
While our ERD model empirically predicts growth via estimating $\mathrm{\Psi}$~leaf~ from $\mathrm{\Psi}$~soil~ of a specific depth, mechanistic models that account for plant hydrodynamics and other processes influencing growth are likely to predict growth more accurately, and thus ERD and hydraulic risk [@Christoffersen:2016kd2d; @Sperry.1998; @Yang.2019; @Duursma:2018]. Although data needs for parameterizing such models could be greater (e.g. vulnerability curves and capacitance for roots, stems and leaves), adding a degree of parameter uncertainty in community-wide application, such models hold greater promise in improving our understanding of plant physiology [e.g. see @Johnson:2018fad].

\
 To estimate topographic variation in ERD within or across species, future studies may use soil moisture dynamics from a distributed hydrological model [e.g. @Schwantes:2018]. We note that data availability on soil water retention curves ($\mathrm{\Psi}$~soil~ vs. VWC), hydraulic conductivity (K~soil~) by depth, soil moisture by depth, stream discharge and evapotranspiration were important for effective calibration of our 1-D hydrological model. We recommend wide-spread and coordinated collection of these variables as well as water-table levels in forest-inventory sites to allow for estimation of trees' water environments.

## Conclusions

Establishing relationships between environment, traits and demographic outcomes of plants is imperative for developing a predictive plant ecology. Trees rooting depths and actual water environments through hydrological rather than meteorological droughts are nonetheless rarely studied. To our knowledge this is the first study to test for a mechanistic link between plant-available water in the whole-soil column, trees' above- and below-ground hydraulic architecture and long-term mortality outcomes for a species-rich forest. We report here that deep-water access plays a role in mitigating mortality of otherwise vulnerable stem hydraulics. This has important implications for our predictive understanding of tropical forest dynamics under current and future climate. Our community-scale framework for modeling effective rooting depths and leaf vulnerability curves indicates the possibilities in expanding the use of these critical, rare observations in species-rich forests towards community-scale generalizations.

# Acknowledgements

This research was supported as part of the Next Generation Ecosystem Experiments-Tropics, funded by the U.S. Department of Energy, Office of Science, Office of Biological and Environmental Research. Funding was also provided to RCT and SMM through an NSF grant (1137366). LR was deputed to Indo-French Cell for Water Sciences on an IRD Fellowship. An NSF grant to Smithsonian Institution’s ForestGEO (Dimensions 1046113) partially supported this work via RCT and SMM. RF was supported by the National Center for Atmospheric Research, which is funded by the National Science Foundation. The BCI forest dynamics research project, founded by Stephen P. Hubbell and Robin B. Foster and sustained for many years by Richard Condit, is now managed by Rolando Perez, Suzanne Lao and Stuart Davies under the ForestGEO program of the Smithsonian Tropical Research in Panama. Numerous organizations have provided funding, principally the US National Science Foundation, and hundreds of field workers have contributed.

# Author contribution

Conceptualization: RCT. Data curation: BF, SP for climate drivers. SP for stream discharge. MTT for GPP, ET, TDR, bulk density. SK for manual GWC and GWC vs $\mathrm{\Psi}_\textrm{soil}$. NK, JZ, KAT, LS for leaf hydraulics, BW for stem hydraulics. RP and SA for the BCI 50ha plot censuses. JW for maximum tree height, WSG, LMA, deciduousness, leaf fall and leaf lifetime. Formal Analysis: RCT, CX. MTT for upgrading GPP-VPD relationship and LAI seasonality. Funding acquisition: SMM, JC, CX, NM, LK. Investigation: RCT, CX (SMM and LR for an exploratory version). Methodology: RCT, CX (SMM and LR for an exploratory version). Project administration: RCT, Resources: SMM, BF, SP, MTT, SK, NK, JZ, KAT, LS, BW, JW. Software: RCT; RF, RK, CF, CX, RCT for ELM-FATES. Supervision: CX, SMM, BN, Validation: RCT. Visualization: RCT, Writing – original draft: RCT. Writing – review & editing: All authors. See https://casrai.org/credit/ for the taxonomy of credits.

# Data Availability

ELM-FATES source code, simulation outputs for best-fit parameter ensemble members and all the R scripts to reproduce the manuscript are available [@Chitra-Tarak.2020]. Data sources for all other datasets used are provided throughout the manuscript.

# References

# Figure Captions 
\
Figure \@ref(fig:schematic): A schematic for the methods workflow.

\
Figure \@ref(fig:hydro-calib): ELM-FATES calibration and evaluation. Observations (red lines and points) versus simulations (gray lines) from 100 best-fit ensemble-member runs of ELM-FATES are shown for daily Volumetric Water Content (VWC) by depth for horizontal time domain reflectometry (TDR) probes at three depths (0.1, 0.4 and 1 m; **a**) and for an average of three vertical TDR probes (0-0.15 m; **b**); monthly stream discharge (Discharge, **c**) and monthly evapotranspiration from eddy flux tower (ET, **d**). Red points in panel **a** show average, manual plot-wide observations of VWC with 95% CI (error bars). All observations (red lines and points) were in-sample, except for VWC data from the horizontal probes (red lines in **a**) which were out-of-sample. Values in inset are average RMSE across the 100 best-fit simulations. For panel **a**, values at the top are for manual, plot-wide VWC and those at the bottom are for VWC from TDR probes. VWC is in units of cm^3^ cm^-3^, while ET and Discharge are in mm month^-1^.

\
Figure \@ref(fig:swp-droughts): Occurrence of extreme hydrological droughts by census intervals (horizontal panels) at three representative depths: 0.1, 0.6 and 1 m (vertical panels). Mean (black lines) and lower half of 95% distribution (grey areas) of soil water potential $\mathrm{\Psi}_{\textrm{soil},z}$ for a given DOY and depth $z$ are shown (same across all census intervals). $\mathrm{\Psi}_{\textrm{soil},z}$ by DOY for a year (colored line) is only shown if at least one day of the year $\mathrm{\Psi}_{\textrm{soil},z}$ was more negative than the 5^th^ percentile of $\mathrm{\Psi}_{\textrm{soil},z}$ for that day of year (DOY).

\
Figure \@ref(fig:ab-fit): Parameters $A$ (panel **a**) & $B$ (panel **b**) for the leaf vulnerability curves that were fitted to observed data on $K_\textrm{leaf}$ vs. $\mathrm{\Psi}$~leaf~ for `r length(sp.exp.param$sp)` species (Eq. \@ref(eq:vc-exp)) vs. those that were predicted from a set of models based on trait-proxies. (Eqs. \@ref(eq:a-mod)-\@ref(eq:b-mod)). Goodness-of-fit ($R$^2^) and significance levels are given in inset.

\
Figure \@ref(fig:erd-iso): Modeled effective rooting depth (ERD; mean $\pm$ 1SE; m) versus dry-season stable isotopic concentration $\mathrm{\delta}$^2^H~xylem~ (mean $\pm$ 1SE; ‰) for `r length(unique(erd.iso$sp))` canopy species from Barro Colorado Island. $\mathrm{\delta}$^2^H~xylem~ data are from @Meinzer:1999.

\
Figure \@ref(fig:sp-erd): Modeled effective rooting depth (ERD; mean $\pm$ 1SE in m) for `r  length(erd.sp)` large, canopy species of Barro Colorado Island. A species ERD is defined as the median, across 100 hydrological realizations, of soil layer depth $z$ at which soil water dynamics ($\mathrm{\Psi}_{\textrm{soil,} z}$) best explained observed dynamics of species growth for each realization. As soil layers in the hydrological model are discretely resolved into exponentially increasing depths, so is the ERD axis. Deeper depths are shown with darker blue colors.

\
Figure \@ref(fig:erd-hyd): Modeled effective rooting depth (ERD) versus hydraulic properties for seven canopy species found on Barro Colorado Island (Panama). Spearman's $r$ and significance levels. Linear model fits (blue lines) with confidence bands (gray area) are shown for significant fits at $\alpha$ = 0.05. See Fig. S\@ref(fig:erd-hyd-full) for the full correlation matrix.

\
Figure \@ref(fig:erd-mort-census-evg): Mortality rate (mean $\pm$ 1SE in % yr^-1^) versus modeled effective rooting depth (ERD; mean $\pm$ 1SE in m) for `r nrow(mrate.depth.mean %>% subset(deciduous == "E"))` evergreen, canopy species found on Barro Colorado Island over seven census intervals (1981-2015). $R$^2^ and significance levels for linear model fits are given in panel insets. Model fits (blue lines) with confidence bands (gray area) are only shown for periods with significant fits at $\alpha$ = 0.1, identified by census-intervals with a superscript *.

\
Figure \@ref(fig:erd-time): Modeled effective rooting depth (ERD; horizontal-axis) versus time spent beyond critical hydraulic threshold (vertical axis) by census interval (colored bars). Each bar represents the average time species of the same ERD spent beyond species-specific critical hydraulic thresholds in a given interval, that is, the proportion of days for which $\mathrm{\Psi}_{\textrm{soil},z=\textrm{ERD}}$ was more negative than $\mathrm{\Psi}$~crit~, defined as $\mathrm{\Psi}$~50,leaf~, and where $z$ is the soil depth matching species ERD. Standard error of the mean are shown over each bar. Data includes all the `r length(erd.mrate.sp)` canopy tree species on Barro Colorado Island that were included in the mortality analyses. Trends were similar for evergreen species alone.

# Figures

```{r schematic, eval = TRUE, echo = FALSE, message = FALSE, fig.align = 'center', out.width = "80%", fig.pos = 'H', results = "asis", fig.cap = paste0("\\label{lab-schematic}")}
knitr::include_graphics(file.path("schematic4/schematic4.001.jpeg"))
```

```{r hydro-calib, eval = TRUE, echo = FALSE, message = FALSE, fig.align ='center', out.width = "80%", fig.pos = 'H', results = "asis", fig.cap = paste0("\\label{lab-hydro-calib}")}
knitr::include_graphics(file.path("figures/hydro-calib2.jpeg"))
```

```{r swp-droughts, eval = TRUE, echo = FALSE, message = FALSE, fig.align ='center', out.width = "100%", fig.pos = 'H', results = "asis", fig.cap = "\\label{lab-swp-droughts}"}
knitr::include_graphics(file.path("figures/PhenoDemoTraitsPsi/psi_model_daily_bestfit_params.top.few_CI_full_interval_panels_climatology_over_study_period_q5_by_depth.jpeg"))
```

```{r ab-fit, eval = TRUE, echo = FALSE, message = FALSE, fig.align ='center', dpi = 96, out.width = "60%", fig.pos = 'H', results = "asis", fig.cap = paste0("\\label{lab-ab-fit}")}
knitr::include_graphics(file.path("figures/PhenoDemoTraitsPsi/kmax_by_psi/Leaf/A_B_data_vs_model.jpeg"))
```

```{r erd-iso, eval = TRUE, echo = FALSE, message = FALSE, fig.align = 'center', out.width = "80%", fig.pos = 'H', results = "asis", fig.cap = paste0("\\label{lab-erd-iso}")}
knitr::include_graphics(file.path("figures/PhenoDemoTraitsPsi/psi.corr_best.depth_xylem_sap_deltaD_phenology_Meinzer_gr.Psi.VPD.jpeg"))
```

```{r sp-erd, eval = TRUE, echo = FALSE, message = FALSE, fig.align ='center', out.width = "60%", fig.pos = 'H', results = "asis", fig.cap= paste0("\\label{lab-sp-erd}")}
knitr::include_graphics(file.path("figures/PhenoDemoTraitsPsi/ERD_by_sp_large_canopy.jpeg"))
```

```{r erd-hyd, eval = TRUE, echo = FALSE, message = FALSE, fig.align ='center', out.width = "60%", fig.pos = 'H', results = "asis", fig.cap = paste0("\\label{lab-erd-hyd}")}
knitr::include_graphics(file.path("figures/PhenoDemoTraitsPsi/erd.stem.traits.jpeg"))
```

```{r erd-mort-census-evg, eval = TRUE, echo = FALSE, message = FALSE, fig.align ='center', out.width = "100%", fig.pos = 'H', results = "asis", fig.cap = paste0("\\label{lab-erd-mort}")}
knitr::include_graphics(file.path("figures/PhenoDemoTraitsPsi/mortality_by rdi.gr_evergreen.jpeg"))
```

```{r erd-time, eval = TRUE, echo = FALSE, message = FALSE, fig.align ='center', out.width = "40%", fig.pos = 'H', results = "asis", fig.cap = paste0("\\label{lab-erd-time}")}
knitr::include_graphics(file.path("figures/PhenoDemoTraitsPsi/mfac vs. rdi.gr.jpeg"))
```

# Tables

```{r hypo, results = 'asis', echo = FALSE}
knitr::kable(hypo.table, caption = 'Hypotheses for association between Effective Rooting Depth (ERD) and hydraulic traits.', booktabs = TRUE, escape = FALSE, col.names = gsub("[_]", "-", gsub("[.]", " ", names(hypo.table))))
```

#
```{r symbols, results = 'asis', echo = FALSE}
knitr::kable(symbols.table, caption = 'The symbol, definition and units of key traits used in the simulations and analyses.', booktabs = TRUE, escape = FALSE)
```