### 2021-01-20 LAI.Rmd
1. Tightened language 2. Removed end period of LAI time series for which leaf gain data not present 3. Added detrended time series

---
title: "LAI workflow"
author: "Rutuja Chitra-Tarak"
date: "1/15/2021"
output:
      bookdown::html_document2:
        number_sections: no
        theme: flatly
        fig_caption: yes
      bookdown::pdf_document2: default
---
```{r setup, echo = FALSE, include=FALSE}
library(knitr)
opts_knit$set(eval.after = "fig.cap")
knitr::opts_chunk$set(echo = TRUE)
if(!require("pacman", quietly=TRUE)) install.packages("pacman")
suppressPackageStartupMessages(
  pacman::p_load(bookdown, knitr, pander, citr, broom))
options(tinytex.verbose = TRUE)
```

1. Built leaf lifespan (LL) to LMA relationship based on both, PNM and San-Lorenzo (FTS) based on canopy leaves of trees (Figure \@ref(fig:ll-lma)), and used that to gap-fill leaf lifespan data.

```{r ll-lma, eval = TRUE, echo = FALSE, message = FALSE, fig.align ='center', out.width = "50%", results = "asis", fig.cap = paste0("\\label{lab-ll-lma}Species-specific Leaf lifespan vs. LMA observed for canopy leaves at two sites at Panama. Each observation (circle) is color coded by site. A linear model fit (line) and equation with R2 and p-value are shown.")}
knitr::include_graphics(file.path("figures/PhenoDemoTraitsPsi/leaf_cohort/lifespan_by_lma_canopy_sitecolored.jpeg"))
```

```{r ll-lma-by-site, eval = TRUE, echo = FALSE, message = FALSE, fig.align ='center', out.width = "50%", results = "asis", fig.cap = paste0("\\label{lab-ll-lma-by-site} Species-specific Leaf lifespan vs. LMA observed for canopy leaves at two sites, shown separately. A linear model fit (line) and equation with coefficient of determination and p-value are shown separately for each site.")}
knitr::include_graphics(file.path("figures/PhenoDemoTraitsPsi/leaf_cohort/lifespan_by_lma_canopy_two sites.jpeg"))
```

2. Restricted leaf-litter data to the last 7 years for BCI-Poachers and BCI-50ha, converted observed leaf fall data column in mass units (leaf_gm) to mass per unit ground area as follows:  
    Leaf in gm per m2 ground area at a site = Leaf biomass in gm / (Site-specific trap area in m2  x Number of leaves were collected from) --- (Eq 1)  
    In Joe's the dataframe, column "leaf_gm"" represents leaf~biomass~in~gm and column "n_traps" represents the number of traps the leaf_gm observation came from.

3. Removed outliers in leaf_gm_per_m2 as those greater than 5 standard deviation by species

4. Remove a small set of species-site combinations with too few records for that species at that site by visually inspecting data

4. Distributed weekly leaf_gm_per_m2 over the days of the week with interpolation

5. When a species' leaf_gm_per_m2 was recorded at both the sites, obtained an average.

6. Converted leaf_gm_per_m2 to LAI units as   
    leaf_fall_LAI (in m2/m2) = leaf_gm_per_m2 / LMA (in g/m2) --- (Eq 2)  
    LMA units are from Metadata_Lftraits_VariableDefinitionsOnly.doc for data in BCITRAITS_20101220.csv)

7. For each species, prepended its leaf_gm_per_m2 data-frame by adding days equal to the species leaf lifespan, sp.lifespan  
    Then LAI time series is calculated: For t in 1 to c(total no of days in the data frame - leaf.lifespan) ,   
    1. Assuming zero LAI to begin with, leaf gain on day 1 is leaf fall for the future day sp.lifespan. That is,    
    If t == 1,  
    LAI[t] = leaf_fall_LAI[sp.lifespan]  
    1. New leaf gain is added to LAI of the previous day, but leaf fall does not begin yet until the day sp.lifespan:  
    For t > 1, if t < sp.lifespan,   
    LAI[t] <- LAI[t-1] + leaf_fall_LAI[sp.lifespan]  
    1. LEaf fall begins on day sp.lifespan, and is removed  
    For t > 1, if t >= sp.lifespan,  
    LAI[t] <- LAI[t-1] + leaf_fall_LAI[sp.lifespan] - leaf_fall_LAI[t]  
    1. I then removed the prepended dates, that is, for t in 1: sp.lifespan, and thus obtained the LAI time series (Figure \@ref(fig:LAI-ts)).

```{r LAI-ts, eval = TRUE, echo = FALSE, message = FALSE, fig.align ='center', out.width = "90%", results = "asis", fig.cap = paste0("\\label{LAI-ts}LAI time series. LL = species-specific leaf lifespan in days.")}
knitr::include_graphics(file.path("figures/PhenoDemoTraitsPsi/leaf_fall/LAI_ts.jpeg"))
```

8. For some species LAI increases/decreases over time likely due to change in population levels. Therefore detrended the time series by removing the slope (Figure \@ref(fig:LAI-ts)).

```{r LAI-ts-det, eval = TRUE, echo = FALSE, message = FALSE, fig.align ='center', out.width = "90%", results = "asis", fig.cap = paste0("\\label{LAI-ts-det}Detrended LAI time series. LL = species-specific leaf lifespan in days.")}
knitr::include_graphics(file.path("figures/PhenoDemoTraitsPsi/leaf_fall/LAI_ts_detrended.jpeg"))
```

9. I then built a species-specific LAI seasonality curve by taking a mean of LAI by DOY. Then normalized this curve by dividing by the max LAI value in the year (Figure \@ref(fig:LAI-ts-det)).

```{r LAI-ssn, eval = TRUE, echo = FALSE, message = FALSE, fig.align ='center', out.width = "90%", results = "asis", fig.cap = paste0("\\label{LAI-ssn}Mean LAI seasonality curves. LL = species-specific leaf lifespan in days. Species are color coded by deciduousness.")}
knitr::include_graphics(file.path("figures/PhenoDemoTraitsPsi/leaf_fall/LAI.seasonality_mean.jpeg"))
```