### 2021-02-06 LAI.Rmd
Reformatted adn updated

---
title: "LAI Workflow Documentation"
author: "Rutuja Chitra-Tarak"
date: "1/15/2021"
output:
      bookdown::html_document2:
        number_sections: no
        theme: flatly
        fig_caption: yes
      bookdown::pdf_document2: default
---
```{r setup, echo = FALSE, include=FALSE}
library(knitr)
opts_knit$set(eval.after = "fig.cap")
knitr::opts_chunk$set(echo = TRUE)
if(!require("pacman", quietly=TRUE)) install.packages("pacman")
suppressPackageStartupMessages(
  pacman::p_load(bookdown, knitr, pander, citr, broom))
options(tinytex.verbose = TRUE)
```

1. Gap-filled species-level mean leaf lifespan with observations from crane sites (data referred to as "cohort""); if not with genus-level mean; else family-level mean. When neither were available: 
    1. For deciduous species, used mean lifespan at PNM site
    2. For evergreen species, used a relationship of leaf lifespan (LL) with LMA. This LMA-based relationship was based on data at San-Lorenzo (FTS) for canopy leaves and for species with sufficient sample sizes (n_lifespan > 100) (Figure \@ref(fig:ll-lma)).

    Per Joe, BCI climate is more similar to San Lorenzo, while PNM site is much drier. More importantly, the BCI 50-ha plot and San Lorenzo support old-growth forests. PNM supports much younger, secondary forest (age unknown, perhaps 100 years). So, used the leaf lifespan-LMA relationship for San Lorenzo. This was used to predict leaf lifespan data for species without leaf lifespan data based on their LMA, after gap-filling species LMA data:  
    Gap-filled LMALAM_AVD with the following sequence of priority:    
    LMALAM_AVD relationship with LMADISC_AVD > LMALAM_AVD rel. with LMALEAF_AVD > Genus-level mean LMALAM_AVD (gap-filled with LMADISC_AVD/LMALEAF_AVD) > Family-level mean LMALAM_AVD (gap-filled with LMADISC_AVD/LMALEAF_AVD)

```{r ll-lma, eval = TRUE, echo = FALSE, message = FALSE, fig.align ='center', out.width = "50%", results = "asis", fig.cap = paste0("\\label{lab-ll-lma}Species-specific Leaf lifespan vs. LMA observed for canopy leaves at San Lorenzo in Panama. Each observation (circle) is color coded by deciduous class. A linear model fit (line) and equation with R2 and p-value are shown.")}
knitr::include_graphics(file.path("figures/PhenoDemoTraitsPsi/leaf_cohort/lifespan_by_lma_canopy.jpeg"))
```

2. BCI-Poachers data goes back to 1985, while BCI-50ha data begins in 2013. I used this entire record. BCI-Poachers has 59 traps of 0.25 m2 each, while BCI-50ha has 62 traps of 0.5 m2 each. Before sending me the leaf-fall dataset, Joe had removed extra traps at BCI-50ha, which were discontinued after a few weeks of observation. I converted observed leaf fall data column in mass units (leaf_gm) to mass per unit ground area as follows:  
    Leaf in gm per m2 ground area at a site = Leaf biomass in gm / (Site-specific trap area in m2 x Total number of traps used ) --- (Eq 1)  
    In Joe's dataframe, column "leaf_gm"" represents leaf biomass in gm.
    Retained only canopy tree species.

3. Distributed weekly leaf_gm_per_m2 over the days of the week with interpolation using base R's approx function.
```r
For a single species, the data-frame looks like this: 
head(df.sp_site[[1]])
       sp4 census leaf_gm n_traps       date         site site.trap.area leaf_gm_per_m2     sp
30024 ALCC      1     0.2       1 1985-11-25 BCI-Poachers          14.75     0.01355932 alchco
30025 ALCC      3    12.0       1 1985-12-09 BCI-Poachers          14.75     0.81355932 alchco
30026 ALCC      4     0.4       1 1985-12-16 BCI-Poachers          14.75     0.02711864 alchco
30027 ALCC      5     0.6       3 1985-12-23 BCI-Poachers          14.75     0.04067797 alchco
30028 ALCC      6     0.2       1 1985-12-30 BCI-Poachers          14.75     0.01355932 alchco
30029 ALCC      7     0.4       2 1986-01-06 BCI-Poachers          14.75     0.02711864 alchco
                  sp_site   form
30024 alchco_BCI-Poachers Canopy
30025 alchco_BCI-Poachers Canopy
30026 alchco_BCI-Poachers Canopy
30027 alchco_BCI-Poachers Canopy
30028 alchco_BCI-Poachers Canopy
30029 alchco_BCI-Poachers Canopy

## First I gap-fill dates in the weekly record data-frame with the following function:

fill.day.gaps <- function(df) {
  df.1 <- df %>% arrange(date) %>%
    # for the first census assume leaf fall is collected over the past week
    mutate(n.days = c(7, as.numeric(date - lag(date))[-1])) %>%
    #  leaf_fall rate per day for the census interval is stored on the day of the census:
    mutate(leaf_gm_rate = leaf_gm_per_m2/n.days,
           # approx() in the next function will interpolate this forward until the next census,
           # we want it to interpolate this value backward over the past census interval,
           # so just for computation purposes, leading this value
           leaf_gm_rate_lead = lead(leaf_gm_rate, n = 1))
  # Gap-filling dates
  full.date.df <- data.frame(date = seq(from = min(df$date, na.rm = TRUE),
                                        to = max(df$date, na.rm = TRUE), by = 1)) %>%
    mutate(day_number = as.numeric(difftime(date, min(df$date, na.rm = TRUE) - 1)))
  df.2 <- df.1 %>%
    full_join(full.date.df, by = "date") %>% arrange(date, site)
  return(df.2)
}
# To get leaf fall on each day of the interval
leaf.fall.daygaps <- lapply(df.sp_site, fill.day.gaps)

head(leaf.fall.daygaps[[1]])
   sp4 census leaf_gm n_traps       date         site site.trap.area leaf_gm_per_m2     sp             sp_site
1 ALCC      1     0.2       1 1985-11-25 BCI-Poachers          14.75     0.01355932 alchco alchco_BCI-Poachers
2 <NA>     NA      NA      NA 1985-11-26         <NA>             NA             NA   <NA>                <NA>
3 <NA>     NA      NA      NA 1985-11-27         <NA>             NA             NA   <NA>                <NA>
4 <NA>     NA      NA      NA 1985-11-28         <NA>             NA             NA   <NA>                <NA>
5 <NA>     NA      NA      NA 1985-11-29         <NA>             NA             NA   <NA>                <NA>
6 <NA>     NA      NA      NA 1985-11-30         <NA>             NA             NA   <NA>                <NA>
    form n.days leaf_gm_rate leaf_gm_rate_lead day_number
1 Canopy      7  0.001937046        0.05811138          1
2   <NA>     NA           NA                NA          2
3   <NA>     NA           NA                NA          3
4   <NA>     NA           NA                NA          4
5   <NA>     NA           NA                NA          5
6   <NA>     NA           NA                NA          6s

## Interpolating from weekly sums to daily leaf_gm
leaf.interp.approx <- function(df) {
  x <- df$day_number
  y <- df$leaf_gm_rate_lead
  xout <- df$day_number[is.na(df$leaf_gm_rate_lead)]
  # yout <- approx(x, y, xout, method = "linear")
  ## method = "constant" would be more parsimonious
  yout <- approx(x, y, xout, method = "constant")
  df.1 <- df %>%
    left_join(data.frame(day_number = yout$x, leaf_gm.int.raw = yout$y), by = "day_number") %>%
    ## filling interpolation gap on the day of the census
    mutate(leaf_gm.int.raw = ifelse(is.na(leaf_gm_rate), leaf_gm.int.raw, leaf_gm_rate),
           sp = df$sp[1],
           site = df$site[1],
           site.trap.area = df$site.trap.area[1])
  return(df.1)
}

leaf.fall.int <- lapply(lapply(leaf.fall.daygaps, leaf.interp.approx),
                        as.data.frame) %>%
  bind_rows(.id = "sp_site") %>%
  select(-leaf_gm_rate_lead)

head(leaf.fall.int)
sp4 census leaf_gm n_traps       date         site site.trap.area leaf_gm_per_m2     sp             sp_site
1 ALCC      1     0.2       1 1985-11-25 BCI-Poachers          14.75     0.01355932 alchco alchco_BCI-Poachers
2 <NA>     NA      NA      NA 1985-11-26 BCI-Poachers          14.75             NA alchco alchco_BCI-Poachers
3 <NA>     NA      NA      NA 1985-11-27 BCI-Poachers          14.75             NA alchco alchco_BCI-Poachers
4 <NA>     NA      NA      NA 1985-11-28 BCI-Poachers          14.75             NA alchco alchco_BCI-Poachers
5 <NA>     NA      NA      NA 1985-11-29 BCI-Poachers          14.75             NA alchco alchco_BCI-Poachers
6 <NA>     NA      NA      NA 1985-11-30 BCI-Poachers          14.75             NA alchco alchco_BCI-Poachers
    form n.days leaf_gm_rate day_number leaf_gm.int.raw
1 Canopy      7  0.001937046          1     0.001937046
2   <NA>     NA           NA          2     0.058111380
3   <NA>     NA           NA          3     0.058111380
4   <NA>     NA           NA          4     0.058111380
5   <NA>     NA           NA          5     0.058111380
6   <NA>     NA           NA          6     0.058111380
  ```

4. When a species' leaf_gm_per_m2 was recorded at both the sites, obtained an average wieghted by the total trap area to accountfor differential sample sizes

5. Converted leaf_gm_per_m2 to LAI units as   
    leaf_fall_LAI (in m2/m2) = leaf_gm_per_m2 / LMA (in g/m2) --- (Eq 2)  
    LMA units are from Metadata_Lftraits_VariableDefinitionsOnly.doc for data in BCITRAITS_20101220.csv)

6. Joe's suggestion: Estimate living canopy leaf mass per unit ground area present at time t (L[t]) from mean leaf longevity and leaf mass fall per unit ground area at time t (F[t]) as follows:  L[t] = F[t+1] + F[t+2] + … + F[t + mean leaf longevity]

    1. For each species, calculated a day of the year average of leaf_fall_LAI (m2/m2), weighted by total trap area of sites observed, to account for differential sampling intensity over time. This gives more weight to dates after 2013 when BCI-50ha site observations began. (Figure \@ref(fig:leaf-fall-LAI))) 
    2. LAI, for the changing fraction on the tree is obtained (Figure \@ref(fig:LAI-ssn-loop)) by   
    looping over days t in 1:365   
    LAI[t] <- sum(leaf_fall_LAI[c(t: (t+sp.lifespan)) %% 365])  
    where sp.lifespan are the number of days in species leaf lifespan  
    To normalise LAI, it is the divided by the maximum value.

```{r leaf-fall-LAI, eval = TRUE, echo = FALSE, message = FALSE, fig.align ='center', out.width = "90%", results = "asis", fig.cap = paste0("\\label{leaf-fall-LAI}Seasonality curves for Leaf Fall in LAI units. LL = species-specific leaf lifespan in days. Species are color coded by deciduousness.")}
knitr::include_graphics(file.path("figures/PhenoDemoTraitsPsi/leaf_fall/Leaf_fall_LAI.seasonality_mean.jpeg"))
```

```{r LAI-ssn-loop, eval = TRUE, echo = FALSE, message = FALSE, fig.align ='center', out.width = "90%", results = "asis", fig.cap = paste0("\\label{LAI-ssn-loop}Mean LAI seasonality curves. LL = species-specific leaf lifespan in days. Species are color coded by deciduousness.")}
knitr::include_graphics(file.path("figures/PhenoDemoTraitsPsi/leaf_fall/LAI.seasonality_mean_loop.jpeg"))
```

```{r LAI-ssn-loop-subcol, eval = TRUE, echo = FALSE, message = FALSE, fig.align ='center', out.width = "90%", results = "asis", fig.cap = paste0("\\label{LAI-ssn-loop-subcol}Mean LAI seasonality curves with species color coded by the source of leaf-lifespan observation/substitution. LL = species-specific leaf lifespan in days.")}
knitr::include_graphics(file.path("figures/PhenoDemoTraitsPsi/leaf_fall/LAI.seasonality_mean_loop_color_substitution.jpeg"))
```