### 2021-04-05 Readme.Rmd
minor

---
title: "Workflow Part II: Inverse root modeling and post-processing"
author: "Rutuja Chitra-Tarak"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document: default
  html_document:
    df_print: paged
  word_document: default
subtitle: Hydraulically-vulnerable trees survive on deep-water access during droughts
  in a tropical forest
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, include = TRUE, eval = FALSE)
```

# Pre-requisite data

**Growth data**

Calculate species-specific growth time series from forest inventory data.

```{r Chunk 01.0, eval = FALSE}
source("code/01.0_Preparing_stem_growth_time_series.R")
```

**Mortality data**

```{r Chunk 02.0, eval = FALSE}
source("code/02.0_demographic_data_BCI.R")
```

#### All pre-requisites

Gather all pre-requisites for rooting depth inverse model (effective rooting depth ) in a place and traits data for post-processing

- Soil Water Potentials (outputs from ELM-FATES)
- VPD-GPP relationship
- Leaf vulnerability curves
- LAI seasonality
- Hydraulic traits

```{r Chunk 05.0, eval = FALSE}
source("code/05.0_Prepare_data_for_correlative analyses.R")
```

Script to load data generated from the data preparation script above for use in inverse model as well as manuscript. It's used within the rest of the scripts.

<!-- ```{r Chunk 06.0, eval = FALSE} -->
<!-- source("code/06.0_load.R") -->
<!-- ``` -->

# Model Setup

Set-up and run effective rooting depth models, evaluate model, and join ERD from the predicted model with hydraulic traits and mortality data, and calculated realized hydraulic risk

```{r Chunk 07.0, eval = FALSE}
source("code/07.0_PhenoDemoTraitsPsi.R")
```

# Manuscript

**Main text**
```{r Chunk 09.0, eval = FALSE}
# Two files are used within the report
# For data, 06.0_load.R and 
# For figures and data, code/08.0_manuscript_data_figs.R
if (!require("groundhog")) install.packages("groundhog"); 
groundhog.folder <- paste0("Users/ftuser/Library/R/4.0/library/bci.roots.inversion")
if(!dir.exists(file.path(groundhog.folder))) {dir.create(file.path(groundhog.folder))}
set.groundhog.folder(groundhog.folder)
library(groundhog)
groundhog.day = "2021-01-01"
# groundhog.library('rmarkdown', groundhog.day)
# .rs.restartR() to restart R session
groundhog.library('bookdown', groundhog.day)
groundhog.library('knitr', groundhog.day)
# rmarkdown::render("09.0_manuscript.rmd", output_format = "bookdown::html_document2")
render("09.0_manuscript.rmd", output_format = "bookdown::word_document2")
```

**Supporting Information**

```{r Chunk 10.0, eval = FALSE}
# render("10.0_Supporting_Information.Rmd", output_format = "bookdown::html_document2")
render("10.0_Supporting_Information.Rmd", output_format = "bookdown::word_document2")
```

```{r Chunk 11.0, eval = FALSE}
remotes::install_github("ropenscilabs/reviewer")
modified_file <- system.file("extdata/CheatSheet-modified.Rmd", package = "reviewer")
reference_file <- system.file("extdata/CheatSheet.Rmd", package = "reviewer")
library(reviewer)
result <- diff_rmd(modified_file, reference_file)
groundhog.library('diffobj', groundhog.day)
browseVignettes("diffobj")
remotes::install_github("ropenscilabs/trackmd")
trackmd::trackChanges("09.0_manuscript.rmd")
```
<!-- groundhog.library('diffr', groundhog.day) -->
<!-- diffr("09.0_manuscript_NP1.rmd", "09.0_manuscript.rmd") -->