### 2021-04-13 09.0_manuscript.rmd
Added a reference

---
title: "Hydraulically-vulnerable trees survive on deep-water access during droughts in a tropical forest"
author: ''
date: ''
output:
  bookdown::word_document2:
        reference_docx: manuscriptstyle.docx
        fig_caption: yes
        number_sections: false
  bookdown::html_document2:
    number_sections: no
    theme: flatly
    fig_caption: yes
  bookdown::pdf_document2: default
bibliography: references.bib
csl: new-phytologist.csl
---

```{r setup1, include=FALSE, echo = FALSE}
if (!require("groundhog")) install.packages("groundhog");
groundhog.folder <- paste0("Users/ftuser/Library/R/4.0/library/bci.roots.inversion")
if(!dir.exists(file.path(groundhog.folder))) {dir.create(file.path(groundhog.folder))}
set.groundhog.folder(groundhog.folder)
library(groundhog)

groundhog.day = "2021-01-01"
pkgs=c('knitr','pander', 'broom')
groundhog.library(pkgs, groundhog.day)

source("code/06.0_load.R")
source("code/08.0_manuscript_data_figs.R")
# load("results/manuWorkSpace.RData")
# install.packages(c("vctrs", "rlang"))
```

```{r setup2, echo = FALSE}
opts_knit$set(eval.after = "fig.cap")
```

Chitra-Tarak, Rutuja*^1,2^, Chonggang Xu^1^, Salomón Aguilar^3^, Kristina Anderson-Teixeira^3,4^, Jeff Chambers^5^, Matteo Detto^3,6^, Boris Faybishenko^5^, Rosie A. Fisher^7,8^, Ryan Knox^5^, Charlie Koven^5^, Lara Kueppers^5,9^, Nobert Kunert^3,4,10^, Stefan J. Kupers^11^, Nathan G. McDowell^12,13^, Brent D. Newman^1^, Steven R. Paton^3^, Rolando Pérez^3^, Laurent Ruiz^14,15,16^, Lawren Sack^17^, Jeffrey M. Warren^18^, Brett T. Wolfe^3,19^, Cynthia Wright^18^, S. Joseph Wright^3^, Joseph Zailaa^4,17,20^, Sean M. McMahon^2,3^

\
*__Corresponding author__: Rutuja Chitra-Tarak  
E: arutuj@gmail.com ;  T: +1 (443) 758 1257

1. Los Alamos National Laboratory, Earth and Environmental Sciences Division (EES-14) 
MS J495, Los Alamos, NM, USA 87545-1663 
2. Smithsonian Environmental Research Center, 647 Contees Wharf Road, Edgewater, MD, USA 21037-0028 
3. Smithsonian Tropical Research Institute, Panama, PA 
4. Conservation Ecology Center, Smithsonian Conservation Biology Institute, Front Royal, VA, USA 
5. Lawrence Berkeley National Laboratory, Climate and Ecosystem Sciences Division, Berkeley, CA, USA 
6. Princeton University, Ecology and Evolutionary Biology, Princeton, NJ, USA 
7. National Center for Atmospheric Research, Climate and Global Dynamics Division
Boulder, CO, USA 
8. Laboratoire Évolution & Diversité Biologique, CNRS:UMR 5174, Université Paul Sabatier, Toulouse, France.
9. University of California Berkeley, Energy and Resources Group, 310 Barrows Hall #3050
Berkeley, CA, USA 94720 
10. Institute of Botany, University of Natural Resources and Life Sciences Vienna, Department of Integrative Biology and Biodiversity Research, Gregor‐Mendel‐Str 33, A‐1190 Wien, Wien, AT 
11. German Centre for Integrative Biodiversity Research (iDiv) Halle-Jena-Leipzig, Computational Forest Ecology, Leipzig, Saxony, DE 
12. Atmospheric Sciences and Global Change Division, Pacific Northwest National Lab, PO Box 999, Richland, WA, USA 99352 
13. School of Biological Sciences, Washington State University, PO Box 644236, Pullman, WA, USA 99164-4236 
14. Indo-French Cell for Water Sciences, Indian Institute of Science, Bangalore, Karnataka, IN 
15. IRD, CNRS, UPS, UMR GET, Toulouse, FR 
16. INRAE, Institut Agro, UMR SAS, Rennes, FR 
17. University of California Los Angeles, Ecology and Evolutionary Biology, 612 Charles E. Young Drive South, Los Angeles, CA, USA 90095 
18. Oak Ridge National Laboratory, Environmental Sciences Division, Oak Ridge, TN, USA 
19. School of Renewable Natural Resources, Louisiana State University Agricultural Center, Baton Rouge, LA 70803, USA 
20. California State University Los Angeles, Biological Sciences Department, Los Angeles, CA, USA

\

Section  | Word Count  
------------- | -------------  
Summary | 200  
Main Text  | 6933  
Introduction | 831    
Methods |  3150  
Results | 838  
Discussion | 2115  
Keywords | 8

\

Item | Count  
------------- | -------------  
Figures in color | 9  
Tables | 2  
Supplementary Material |  1 file

# Summary

- Deep-water access is arguably the most effective, but under-studied mechanism that plants employ to survive during drought. Vulnerability to embolism and hydraulic safety margins can predict mortality risk at given levels of dehydration, but deep-water access may delay plant dehydration. Here, we tested the role of deep-water access in enabling survival within a diverse tropical forest community in Panama using a novel data-model approach.
- We inversely estimated the effective rooting depth (ERD, as the average depth of water extraction), for 32 canopy species by linking diameter growth dynamics (1990-2015) to vapor pressure deficit, water potentials in the whole-soil column, and leaf hydraulic vulnerability curves. We validated ERD estimates against existing isotopic data of potential water-access depths.
- Across species, deeper ERD was associated with higher maximum stem hydraulic conductivity, greater vulnerability to xylem embolism, narrower safety margins, and lower mortality rates during extreme droughts over 35 years (1981-2015), especially in evergreen species. Species exposure to water-stress declined with deeper ERD indicating that trees compensate for water-stress related mortality risk through deep-water access.
- The role of deep-water access in mitigating mortality of hydraulically vulnerable trees has important implications for our predictive understanding of forest dynamics under current and future climates.
\
Keywords: rooting depths, deep-water access, hydraulic vulnerability and safety margins, drought-induced mortality, drought tolerance, tropical forest, safety-efficiency trade-off, hydrological droughts

# Introduction

Drought-induced mortality in tropical forests may have significant global implications. Tropical forests play a disproportionately large role in the global carbon and energy cycles [@Bonan:2008], and support half of global biodiversity [@Wright:20050th], but face a threat from intensifying droughts [@Malhi.2008; @Doughty:2015; @Xu.2019]. Tropical forests are considered an especially drought-vulnerable biome given the combination of climate risk and vegetation sensitivity [@Meir:2015trw]. Still, mortality events in the tropics are rarely as large as those in temperate and boreal zones [@McDowell.2018], leading to questions regarding the role of large trait-diversity and hydraulic strategies in mitigating mortality events. Further, mortality rates are increasing in some tropical regions [@Brienen.2015; @Hubau.2020] and widespread drought-induced tree mortality has occurred across the tropics for specific functional groups [@Bennett:2015; @Chitra-Tarak.2018; @Hilker:2014; @Phillips:2010]. State-of-the-art dynamic global vegetation models (DGVMs) struggle to capture these drought-induced vegetation dynamics in the tropics [@Galbraith.2010; @Powell:2013; @Powell.2018], because underlying mechanisms of drought-tolerance are not fully understood nor quantified.

\
Plants rely on a variety of structural and functional mechanisms to avoid or tolerate a drought, from deep-water access, increased root production, hydraulic redistribution, embolism resistance, adjustment of leaf area (deciduousness) to change in leaf angle, reductions in stomatal conductance, upregulation of aquaporins, osmotic regulation and stem water storage capacitance  [@McDowell.2008]. Deep-water access is arguably the most effective, yet under-studied mechanism. As plant-available water varies with depth, trees within the same forest with different rooting depths, depending on species and size [@Meinzer:1999; @Chitra-Tarak.2018; @Brum.2019], differ in their experience during a drought, and thus in their growth and mortality responses [@Chitra-Tarak.2018]. A key bottleneck in community-wide testing of this mechanism has been a lack of data in both trees' rooting or water-sourcing depths and plant-available soil water at those depths. A recent meta-analysis documented maximum rooting depths for 318 tree species, that is, less than 0.5% of more than 60 thousand tree species in the World [@Fan:2017; @Beech:2017]. Further, only a small fraction of those are tropical, even though more than 90% of the World’s tree diversity resides in the tropics [@Slik:2015]. The use of stable isotopes of water as a tracer provides an indirect measure of water sourcing depths by matching the isotopic value in xylem water to those in soil pore water at different depths. However, such data are rare [@Evaristo:2016]. DNA barcoding of roots [@Jones.2011] may be used to estimate species-specific rooting depths or profiles, but DNA barcode libraries for tropical forests are still under development, and the method may be cost-prohibitive for the extent of sampling required. In general, community-scale data collection for rooting or water-sourcing depths is a formidable challenge in species-rich tropical forests.

\
Characterizing the essential constraints to model species-rich communities [@Koven:2020; @Christoffersen:2016kd2d; @Maréchaux.2017; @Wright:2010; @Bartlett.2019; @Lu.2020] entails identifying the topography of plant trait trade-offs in different environments, and how these relate to demographic rates (growth, recruitment and mortality). Significant efforts have been invested into identifying and linking universal drought indices, above-ground traits, and demographic rates. Such efforts have found correlations between vulnerability to embolism and hydraulic safety margins and mortality [@Anderegg:2016], although numerous counter examples also exist [@Hoffmann.2011; @Johnson:2018fad; @Nardini:2015; @Paddock:2013; @Venturas.2016]. Deep roots may mitigate hydraulic vulnerability [@Brum.2019] and mortality risk, in particular during hydrological (rather than meteorological) droughts. Nonetheless, the interaction between rooting depths, above-ground hydraulic traits, hydrological droughts quantified over the whole soil-column and mortality outcomes are hardly studied.

\
In this paper, we estimate plant-available water in the whole-soil column in a tropical forest, and inversely estimate rooting depths of co-occurring tree species from their growth responses, with a series of model calibrations and validation. We evaluate how rooting depth is linked to above-ground hydraulic traits and mortality rates through seven census-intervals over a 35-year period that experienced El Niño droughts of a variety of intensity, frequency and duration [@Detto.2018; @Condit.2017]. Our hypotheses are that (i) deep-rooted trees have hydraulic traits associated with rapid water transport but cavitation-vulnerable xylem due to greater and more reliable water availability at depth (see Tables \@ref(tab:symbols) and \@ref(tab:hypo)); and that, (ii) deep-rooted species have lower mortality rates during droughts due to a lower exposure to water-stress compared to shallow-rooted species. To our knowledge this is the first study to test for a mechanistic link between plant available water in the whole-soil column, trees' above- and below-ground hydraulic traits and multi-decadal mortality outcomes for a species-rich tropical forest.

# Materials and Methods

This work combines hydrology, physiology and demography of a tropical forest at Barro Colorado Island, Panama, to inform the inverse model for rooting depths; validates the model, and tests hypotheses pertaining to relationship of rooting depths with above-ground hydraulic traits, drought exposure, and mortality (Fig. \@ref(fig:schematic)). We defined species-specific effective rooting depths (ERD) as the depth at which the growth factor determined by soil water potential and leaf hydraulic traits best explained species' growth dynamics over 25 years. Developing a novel, empirical inverse model, we estimated ERDs for large trees of `r length(erd.sp)` species. The ERD model incorporated the impact of atmospheric and hydrological drought on growth, and was constrained with species-specific leaf vulnerability curves. For the latter, we used existing data for eight species and developed trait-based proxies for the rest (based on data for a total of 21 species). We obtained the daily dynamics of soil water potential in the whole-soil column (up to 13 m) over the 25 years by locally parameterizing a 1-D hydrological water balance of the forest within a land surface model for an average vegetation type. The water balance was calibrated on available measurements, in particular continuous soil moisture data in three surface layers (over the first meter of soil), stream discharge and evapo-transpiration.

\
We validated our ERD estimates against existing stable hydrogen isotope ratios ($\mathrm{\delta}$^2^H~xylem~) for a subset of six tree species as independent observations. We evaluated whether ERD was associated with above-ground hydraulic traits sourcing the latter from a set of rare datasets for six to seven species that overlapped with our assessment of ERD. We tested the relationship of ERD with xylem vulnerability to embolism, branch hydraulic safety margins, leaf turgor loss point and maximum stem-specific hydraulic conductivity. We analysed whether ERD is correlated with mortality dynamics of large trees for evergreen and deciduous species over 35 years marked by several El Niño events. Finally, to test whether ERD explained risk of drought-induced mortality, we analysed species-level exposure to water-stress.

## Study site description

We conducted this study at Barro Colorado Island (BCI), Panama. The entire island is forested and classified as a tropical moist forest in the Holdridge Life Zone system. Long-term hydrological monitoring at BCI began in 1972, while demographical monitoring began with the 50 ha ForestGEO plot establishment in 1981-82 [@Condit.1998uq; @Hubbell:1999; @Anderson-Teixeira:2014]. Rainfall at BCI is seasonal with a mean annual total of `r round(rain.man.stats$rain)` mm ($\pm$ `r round(rain.man.stats$rain.sd)` SD; `r min(clim$Year)`-`r max(clim$Year)`) and a pronounced dry season from mid-December through April with less than 100 mm of rainfall per month [@Paton.2020; Paton@2019]. In the 50ha (1000 m × 500 m) old growth forest plot, all stems $\ge$ 1 cm diameter at breast height (dbh) are mapped, tagged with a unique number, identified to species and measured every five year through 1985-2015 for growth, mortality as well as recruitment of new stems into the 1 cm dbh size class [@Condit.2017]. This inventory represents 321 woody species, 28% of which are at least partly dry-season deciduous [@Condit:2000]. The elevation range in the plot is only 40 m (120-160 m) [@Harms:2001] and soils are homogeneous with red light clays accounting for 72% of the plot [@Baillie.2017]. The topsoil field texture is silty clay loam that gradually fines to silty clay in the subsoil. Soil is mostly free draining, but restricted subsoil permeability gives rise to temporary wet season ponding. Detailed descriptions of the climate, geology, flora and fauna of BCI can be found elsewhere [@Croat.1978; @Leigh.1982; @Gentry.1990].

## Model for effective rooting depth (ERD)

### ERD model description

Roots could impact tree growth through two factors: 1) water uptake; and 2) nutrient uptake. Because our study is situated in an old-growth tropical forest, we assume that nutrients are mainly concentrated in the shallow soil layers derived from litter decomposition, and that rooting depths do not substantially affect nutrient uptake. Because our study site has a dry season, we expect that rooting depth is a key factor affecting tree growth. We defined species-specific effective rooting depth (ERD) as the depth at which a soil moisture growth limitation factor, $B$, determined by the soil water potential and hydraulic traits best explained species’ diameter growth dynamics. $B$ is used to approximate the amount of stomatal closure due to water-stress in the soil. In this study, we used the fractional loss of hydraulic conductivity of the leaf ($\textrm{FLC}$~leaf~) to estimate $B$. $\textrm{FLC}$~leaf~ is calculated as a fraction of maximum leaf conductivity $K_\textrm{max, leaf}$,

\begin{equation}
\textrm{FLC}_{\textrm{leaf}} = K_{\textrm{leaf}}/K_{\textrm{max, leaf}}
(\#eq:flc)
\end{equation}

where $K$~leaf~ is the current leaf hydraulic conductance and $K_\textrm{max, leaf}$ is the maximum leaf hydraulic conductance (see Table \@ref(tab:symbols) for trait symbols and their definitions). The dynamics of $K$~leaf~ are estimated from the species-specific relationship between $K$~leaf~ vs. $\mathrm{\Psi}$~leaf~, referred to as the leaf hydraulic vulnerability curve [@Sack.2012]. The vulnerability to loss of hydraulic conductivity arises not only from embolism but also from extra-xylem processes [@Scoffoni:2017]. Because pre-dawn $\mathrm{\Psi}$~leaf~ generally approaches $\mathrm{\Psi}$~soil~, this allowed us to substitute $\mathrm{\Psi}$~leaf~ with $\mathrm{\Psi}$~soil~ in leaf hydraulic vulnerability curves for each species $s$---defined below in Eq. \@ref(eq:vc-exp) using species-specific parameters $A_{s}$ and $B_{s}$---to predict maximum diurnal $K_{\textrm{leaf},s, i}$ for each day $i$,

\begin{equation}
K_{\textrm{leaf},s, i} = A_{s}e^{-B_{s}{\mathrm{\Psi}}_{\textrm{soil, i}}}
(\#eq:k-soil)
\end{equation}
$K_{\textrm{max, leaf}, s}$ was obtained as the maximum value of $K_{\textrm{leaf}, s}$ using Eq. \@ref(eq:k-soil). $K$~leaf~ is strongly related to photosynthetic capacity [@Brodribb.2002]. Because we were interested in relating $K$~leaf~ to 5-year average diameter growth observations, we ignored diurnal dynamics of $K$~leaf~.

\
Many other intrinsic and extrinsic factors limit plant growth, including soil moisture, VPD, radiation, leaf area seasonality [@Lawrence.2019; @Brodribb.2002]. VPD affects growth non-linearly, with growth increasing with VPD up-to a VPD threshold, then decreasing as leaf pores (stomata) close reducing water-uptake [@Grossiord.2020; @Yang.2019]. To account for the non-linear impact of VPD on growth, we used predicted gross primary productivity (GPP; hereafter, $\hat{\textrm{VPD}}$) from a locally derived polynomial relationship between GPP and VPD (see SI and Fig. S1). Apart from stomatal control, leaf deciduousness may further limit water-uptake and growth.

\
We therefore tested alternate structures of empirical growth models (Eq. \@ref(eq:M-chosen) and those in SI), or inverse models of ERD, in which we regressed species-specific growth against multiplicative or additive effects of one or more growth factors calculated daily and averaged over 5-yearly censuses: $\hat{\textrm{VPD}}$, $\textrm{FLC}_\textrm{leaf}$ and the leaf area index (LAI). Incorporation of radiation did not improve model fitting, possibly due to the coincidence of higher temperature, higher radiation, and lower humidity during the dry season at BCI.

\
The best empirical growth model, or inverse ERD model, structure that we found (see model validation and selection below) to describe daily average tree growth $\hat{G}$ for species $s$ in the census interval $t$ is described as follows, in which $K_\textrm{leaf}$, and thus $FLC_\textrm{leaf}$, is driven by soil water dynamics at $z$:

\begin{equation}
\hat{G}_{s,t|z} = \beta_{0,s|z} + \beta_{1,s|z}(\frac{1}{n_{t}}\sum_{i=1}^{n_{t}} \textrm{FLC}_{\textrm{leaf},s,i|z}^{*} \hat{\textrm{VPD}}_{i}^{*}) + \epsilon_{s,t|z}
(\#eq:M-chosen)
\end{equation}
where | refers to conditionals, n~t~ is the total number of days in census interval $t$, the superscript * indicates that the variable has been standardized to range between 0 and 1 (within species for $\textrm{FLC}$~leaf~), $\beta_{0}$ and $\beta_{1}$ are model coefficients and $\epsilon$ is the model error term. See SI for all the alternate models tested.

\
We evaluated different model structures and for each we estimated species effective rooting depths (ERD) as the depth $z$ at which soil water dynamics ($\mathrm{\Psi}_{\textrm{soil,} z}$) best explained observed dynamics of growth $G$ (see below) via modeled growth $\hat{G}$. Our growth model, or inverse ERD model, does not explicitly use rooting profiles, but identifies soil water dynamics at a single depth $z$ as the central tendency that influences the observed growth dynamics the most. We modelled multiple hydrological realisations of soil water-potential dynamics (see below). Incorporating this uncertainty, we defined species ERD as the median ($\pm$ SE) of best-fit depths across all hydrological realisations. See SI for statistics for identifying best-fit ERD.

### Growth data

For diameter growth estimates, to minimize the effect of light variation among trees, we selected only large trees ($\ge$ 30 cm dbh) in the 50 ha plot and also species whose maximum height was $\ge$ 30 m (hereafter, canopy species) and thus are likely to be fully exposed to the sun. Calculating individual tree growth rates across six 5-yearly censuses (1990-2015; @Condit.2019a; @Condit.2019b), removing outliers, obtaining residuals from a dbh model of growth to account for the size effect on growth (see SI), we estimated species $s$ growth time series $G_{s}$ (cm yr^-1^) as the median of standardized dbh model residuals, for only those species ($n$ = `r length(erd.sp)`) with complete records for at least three trees (median `r median(g.n$n)`, max `r max(g.n$n)` trees per species).

### Leaf hydraulic vulnerability curves

We obtained leaf hydraulic vulnerability curves ($K_\textrm{leaf}$ vs. $\mathrm{\Psi}$~leaf~) for adult trees of 21 common species at BCI from Zailaa *et* *al.* (in prep; see SI for brief description of methods) described as:

\begin{equation}
K_{\textrm{leaf},s} = A_{s}e^{-B_{s}{\mathrm{\Psi}}_{\textrm{leaf}_{s}}} + \epsilon_{s}
(\#eq:vc-exp)
\end{equation}
where $A$ and $B$ are fitted species-specific parameters, $\epsilon_{s}$ an error term. These 21 species included eight of the `r length(erd.sp)` species selected for ERD estimation. For the remaining `r sp.n.model` species, we obtained vulnerability curves using trait-based proxies. We identified scaling relationships between fitted parameters $A$ and $B$ in Eq. \@ref(eq:vc-exp) and two traits; namely, WSG, the wood specific gravity, and LMA, the leaf mass per unit area (@Wright:2010; see SI). We fitted polynomial equations described as,

\begin{equation}
\begin{split}
B_{s} = `r bcf[1]` `r bcf[2]`\textrm{WSG}_{s} + `r bcf[3]`\textrm{WSG}_{s}^2 `r bcf[4]`\textrm{WSG}_{s}\textrm{LMA}_{s} \\
& + `r bcf[5]`\textrm{LMA}_{s} `r bcf[6]`\textrm{LMA}_{s}^2 + \epsilon_{s,b}
(\#eq:b-mod)
\end{split}
\end{equation}

\begin{equation}
A_{s} = `r acf[1]` `r acf[2]`B_{s} `r acf[3]`B_{s}^2 + `r acf[4]`B_{s}\textrm{LMA}_{s} + `r acf[5]`\textrm{LMA}_{s} `r acf.6`\textrm{LMA}_{s}^2 + \epsilon_{s,a}
(\#eq:a-mod)
\end{equation}
where $\epsilon_{s,a}$ and $\epsilon_{s,b}$ are error terms.

\
As these fits explained a large proportion of variation in parameters $A$ and $B$ (see results), we sequentially used Eq. \@ref(eq:b-mod) and \@ref(eq:a-mod) to predict parameters $B$ and $A$, respectively, for the `r sp.n.model` ERD species without direct data and estimated leaf hydraulic vulnerability curves using Eq. \@ref(eq:vc-exp). We thus obtained parameters $A$ and $B$ for all the `r length(erd.sp)` ERD species for use in Eq. \@ref(eq:k-soil). We also estimated species $\mathrm{\Psi}$~20,leaf~ using their vulnerability curves.

### Leaf Area Index (LAI)

In some of the alternative models for effective rooting depth, we explored the effect of seasonality in leaf area index (LAI) on growth. We assumed species-specific mean seasonal curves for LAI (standardised between 0 and 1; unitless), informed by a combination of long-term records for leaf-fall [@Wright.1990aeo] and leaf-lifetime [@Osnas.2018]. See SI.

### Using ELM-FATES to model soil matric potentials

#### ELM-FATES model description

To estimate dynamics of water availability by depth over the forest’s rooting zone, we used the Energy Exascale Earth System Land Model (ELM; @Caldwell.2019), coupled with the Functionally Assembled Terrestrial Ecosystem Simulator (FATES; @Koven:2020) (hereafter, ELM-FATES). ELM is a land-model that, among many features, simulates the physics and conservative dynamics of water, energy and carbon fluxes. In particular, soil hydrologic fluxes are vertically resolved among discrete soil layers (1-D) similarly to the CLM4.5 [@Oleson.2013]. FATES is a community-based, open-source model used for studying climate-vegetation interactions. FATES is a vegetation demography model, with a size-structured group of plants (cohorts) and successional trajectory-based patches based on the ecosystem demography [@Moorcroft.2001] approach. FATES couples to ELM by a common interface of water and carbon fluxes. Detailed descriptions for ELM and FATES can be found elsewhere [@Bisht.2018; @Fisher.2010; @Fisher.2014; @Koven:2020].

\
  We ran ELM with FATES vegetation, in which the ELM model simulates interception, throughfall, canopy drip, infiltration, evaporation, surface runoff, subsurface drainage, redistribution within the soil column, and groundwater discharge and recharge so as to simulate changes in canopy water, surface water, soil water by depth and water in an unconfined aquifer (omitting processes relevant to snow, wetlands or lakes). See SI for a water balance equation and a note on how soil water dynamics is simulated in ELM-FATES. The soil profile is discretized into maximum 15 exponentially distributed soil layers with layer node depth $z$. Here, $z$ $\in$ $Z$; $Z$ = (0.01, 0.03, 0.06, 0.12, 0.21, 0.37, 0.62, 1, 1.7, 2.9, 4.7, 7.8, 13) m.

#### ELM-FATES model parameterization

To parameterize catchment hydrology in ELM-FATES, we identified eleven parameters relevant for the water balance, determined their ranges based on literature for the study site, else for the tropics (Table S1), and ran 5000 simulations using Latin Hypercube Sampling [LHS; @Stein.1987] from this global parameter space. Notably, we leveraged local data for soil hydraulic conductivity by depth (@Godsey:2004; Fig. S2), and instead of the ELM default soil texture-based pedo-transfer functions, we estimated parameters of soil retention curves using existing data for gravimetric water content vs. $\mathrm{\Psi}$~soil~ [@Kupers:2019]. See SI.

\
We ran ELM-FATES for the 5000-member 11 parameter ensemble with hourly climate drivers measured at the BCI meteorological station over 1985-2018 [@Faybishenko.2021] initialized with the observed stand structure from the 50 ha plot [@Condit.2019a] in a single site mode. As our key interest here was on deriving soil water availability, we ran ELM-FATES in a lower-complexity configuration: static stand structure (see SI), and with a single plant functional type (PFT) of evergreen trees. The latter was chosen as only 9.7% of BCI crown area is dry-season deciduous [@Condit:2000] and addition of a dry-deciduous PFT did not significantly alter results (not shown).

#### ELM-FATES model calibration

We calibrated ELM-FATES over 2012-2018 against three key fluxes and states in the water balance equation, namely, (1) evapotranspiration ET from the flux tower by the 50 ha plot (2012-2017), (2) local stream discharge (2012-2018; see @Paton.2019) and (3) soil volumetric water content (VWC) from two sources: (i) a long-term (2012-2018) record of VWC averaged across three vertical time domain reflectometry (TDR) probes over depth 0-15 cm from three locations near the flux tower, and (ii) plot-wide snap-shot measurements of VWC during the dry season of 2015 and 2016 at depths of 0.15, 0.4 and 1 m (1299 samples covering all soil types and habitats; @Kupers:2019).

\
 For ELM-FATES calibration we calculated an objective function for each of the 5000-member ensemble by equally weighting standardized root mean square error (RMSE) between observations and simulations across all fluxes and states mentioned above, and then identified 100 parameter ensemble members that minimised the objective function, ensuring that soil moisture dynamics by depth is captured correctly (see SI). See Table S1 for the ranges of best-fit values for different parameters.

#### Soil Water Potential Dynamics and Hydrological droughts

We ran ELM-FATES with the best-fit 100 ensemble members from 1985-2018, with the first five years used for model spin-up. Extreme hydrological droughts were identified by depth $z$ as days for which $\mathrm{\Psi}_{\textrm{soil},z}$ was more negative than the 5^th^ percentile of $\mathrm{\Psi}_{\textrm{soil},z}$ for a given day of the year.

#### ELM-FATES model evaluation

We evaluated ELM-FATES by calculating RMSE between simulated and observed long-term daily VWC for the depths of 0.1, 0.4 and 1 m (2016-2018), based on a dataset we had left-out during calibration. We obtained these observations from three horizontal TDR probes at the depths of 0.1, 0.4 and 1 m at a location near the vertical probes (see [SI](#SIData.VWC)).

### ERD model structure selection

We used $\mathrm{\delta}$^2^H~xylem~ as an independent observation to validate the ERD models. As root water-uptake is generally a non-fractionating process, tree $\mathrm{\delta}$^2^H~xylem~ reflects the signature of source water. Given a vertical gradient of $\mathrm{\delta}$^2^H in soil and groundwater, $\mathrm{\delta}$^2^H~xylem~ provides an index of rooting depth [@Dawson.1991]. We leveraged $\mathrm{\delta}$^2^H~xylem~ from BCI [@Meinzer:2001] for the dry-season of March 1997 as this period showed largest seasonal divergence in $\mathrm{\delta}$^2^H~xylem~ among species and vertically in soil and groundwater $\mathrm{\delta}$^2^H at natural abundance level (Fig. S3).

\
For comparison with modelled ERD, we removed six species from the @Meinzer:1999 dataset to account for the uncertainty in their water-sourcing depths. $\mathrm{\delta}$^2^H~xylem~ from leafless trees may not be linked with water sourced at the time of measurement and thus may not be comparable to species that had leaves. Leaflessness status of sampled trees is not recorded in @Meinzer:2001. We therefore removed five species that are typically leafless in March-April (Joseph Wright, personal observation), that is, the months of $\mathrm{\delta}$^2^H~xylem~ sampling. $\mathrm{\delta}^2H$ to depth relationship was particularly uncertain for $\mathrm{\delta}^2H$ > -40 ‰, so from the remaining dataset, we removed one species, *Guapira standleyana*, with $\mathrm{\delta}$^2^H~xylem~ of -28.9 ‰ $\pm$ 3.7SE (see Fig. S3). For each model of ERD (Eq. \@ref(eq:M-chosen) and alternatives in SI), modeled species ERD was regressed against species $\mathrm{\delta}$^2^H~xylem~ for a maximum of six species.

## Relationships between ERD and above-ground hydraulic traits

We evaluated whether ERD was associated with above-ground hydraulic traits sourcing the latter from existing datasets [@Wolfe.2019; @Wolfe.2021] for seven species that overlapped with our assessment of ERD. We regressed species ERD against maximum stem hydraulic conductivity $K_\textrm{max,stem}$ (*n*=7), leaf turgor loss point $\mathrm{\Psi}$~tlp~ (*n*=7), vulnerability to embolism from cavitation $\mathrm{\Psi}$~88,stem~ measured in terms of pressure at which 88% of $K_\textrm{max,stem}$ is lost (*n*=7), and hydraulic safety margins $\mathrm{\Psi}$~min~ - $\mathrm{\Psi}$~88,stem~ (*n*=6). See [SI](#SIData.brett) and Table S2 for data collection and estimation of these variables.

## Mortality analyses and species-specific drought exposure

To test whether ERD plays a role in mitigating mortality risk, we calculated mortality rates for large trees (here, $\ge$ 10 cm dbh) in the 50ha plot for species with ERD estimates (all canopy species with maximum height $\ge$ 30 m) as well as average abundance of at least 20 trees in the plot ($n$ = `r length(erd.mrate.sp)`) [@Condit.2019a]. For each of the seven census intervals $t$ in the 35-yr record (1981, 1985, 1990, 1995, 2000, 2005, 2010, 2015), mortality rate, $\textrm{M}_{t}$ (% yr^-1^) for species $s$ for two successive censuses, $c_{1}$ and $c_{2}$, was calculated as $\textrm{M}_{s,t} = \frac{D_{s,c_{2}}}{N_{s,c_{1}}}\times\frac{100}{\textrm{d}}$, where $N_{s,c_{1}}$ and $D_{s,c_{2}}$ are the total number of large trees of species $s$ present in the 50 ha plot in $c_{1}$ and dead in $c_{2}$, respectively, and $d$ is the duration based on mean dates of $c_{1}$ and $c_{2}$.

\
Dry season deciduous species may escape drought exposure via leaf deciduousness, so we analysed deciduous and evergreen species separately. Tree species on BCI are scored by expert botanists as one among four leaf-habits; evergreen, brevideciduous, facultative deciduous and obligate deciduous (@Meakem.2018; see SI). We pooled all deciduous leaf habits together (hereafter, 'deciduous' for brevity) and regressed species ERD against species $\textrm{M}_{s,t}$ for each census interval $t$ for deciduous ($n$ = `r nrow(mrate.depth.mean %>% subset(deciduous != "E"))`) and evergreen species ($n$ = `r nrow(mrate.depth.mean %>% subset(deciduous == "E"))`) separately.

\
As an indicator of exposure to water-stress, we use a species-specific critical hydraulic threshold, $\mathrm{\Psi}$~crit~, here defined as $\mathrm{\Psi}$~20,leaf~. The duration of exposure to water-stress, and thus the potential for realized hydraulic risk, was defined for each species as the proportion of days in each census interval $t$ of the $\mathrm{\Psi}_\textrm{soil}$ simulation period (1990-2015) during which soil water potential in the soil layer matching species ERD, $\mathrm{\Psi}_{\textrm{soil},z=\textrm{ERD}}$, were more negative than species $\mathrm{\Psi}$~crit~.

\
All statistical analyses were conducted in the *R* statistical environment (version 4.0.3, @R:2020).

# Results

## Soil water dynamics by depth and hydrological droughts

The 100 ELM-FATES ensemble members with the best fits to observed soil moisture dynamics captured soil moisture seasonality at multiple depths (Fig. \@ref(fig:hydro-calib)**a**, **b**), including the out-of-sample observations. These simulations also captured the dynamics in stream discharge (Fig. \@ref(fig:hydro-calib)**c**) and in evapotranspiration (Fig. \@ref(fig:hydro-calib)**d**), slightly underestimating peak discharge in wet years and slightly overestimating peak ET. The reduction in parameter range (Table S1) in the best-fit ensembles compared to the tested global ranges showed that the model calibration was primarily sensitive to the Ball-Berry stomatal slope parameter (fates_leaf_BB_slope), the ELM root distribution parameter that regulates the depth of the rooting profile (fates_rootb_par), soil hydraulic conductivity (HKSAT) profile especially at depth, and the adjustment factor (HKSAT_ADJ) that modifies soil hydraulic conductivity to account for macroporosity and direct flow paths (Figs. S4, S5). The distribution of maximum depth of soil water dynamics, and thus ecosystem root zone depth, across the 100 hydrological realisations encompassed 95% CI of `r root.zone.CI1` m - `r root.zone.CI2` m (median at `r root.zone.med` m).

\
Soil water potential dynamics $\mathrm{\Psi}_{\textrm{soil},z}$ by depth z from 0.01 through 13 m obtained from the best-fit simulations revealed that every 5-yr census interval had at least one extreme hydrological drought year, but each interval varied in terms of number of extreme years, drought intensity, drought seasonality and duration (Fig. \@ref(fig:swp-droughts)). Hydrological droughts in census interval 1990-1995 was marked by a prolonged dry season, while those during 2000-2005 and 2005-2010 distinctly occurred in the wet season, effectively extending the dry season into the wet season (Fig. \@ref(fig:swp-droughts)). Across the simulation period 1990-2015, simulated $\mathrm{\Psi}_{\textrm{soil},z}$ remained above -0.5 MPa from depths 1.7 to 13 m (Fig. S6).

## Predictors of leaf hydraulic vulnerability curves

Parameters of leaf vulnerability curves were predictable from WSG and LMA (Fig. \@ref(fig:ab-fit)). WSG and LMA explained a large proportion of variance in parameter $B$ (Eq. \@ref(eq:b-mod), Adj. $R$^2^ = `r bfit.r2`, $p$ `r  bfit.p`). WSG and parameter $B$ explained a large proportion of variance in parameter $A$ (Eq. \@ref(eq:a-mod), *Adj. R*^2^ = `r afit.r2`, $p$ `r  afit.p`; also see Fig. S7). This predictive power allowed us to estimate leaf vulnerability curves used in the ERD models for `r sp.n.model` ERD species that lacked direct observations (among `r length(erd.sp)` ERD species) (Table S3, Fig. S8).

## Effective rooting depths (ERD)

The best ERD model (Eq. (3)) explained a large fraction of the variance in $\mathrm{\delta}$^2^H~xylem~ ($R$^2^ = `r lm.erd.iso.r2`, $p$ = `r lm.erd.iso.p`, $n$ = `r length(unique(erd.iso$sp))`, Fig. \@ref(fig:erd-iso); also see SI). This model included an effect of VPD and not LAI (Eq. \@ref(eq:M-chosen)).

\
Modeled effective rooting depths (ERD) for the `r length(erd.sp)` large ($\ge$ 30 cm DBH) trees of canopy species varied from `r min(df.erd.to.plot$depth)` m to `r max(df.erd.to.plot$depth)` m (Fig. \@ref(fig:sp-erd)). Evergreen and deciduous species had similar range of ERDs, but a greater proportion of deeper ERD species tended to be evergreen rather than deciduous---a group composed of a variety of categories of deciduousness (Fig. \@ref(fig:sp-erd)). Notably, two species, *Luehea seemannii* and *Trichilia tuberculata*, that @Meinzer:1999 found to have $\mathrm{\delta}$^2^H~xylem~ values between soil water and groundwater, suggesting that these species sourced most of the water from depths >1 m and likely to have sourced some portion of ground water, were also identified by our model with ERD >1 m (2.9 m for both species, Fig. \@ref(fig:erd-iso)).

## Relationships between ERD and above-ground hydraulic traits

Species with deeper ERD showed greater $K_\textrm{max,stem}$ (Spearman's' $r$ = `r cor.k`, $p$ = `r p.k`, Fig. \@ref(fig:erd-hyd)**a**), less negative leaf $\mathrm{\Psi}$~tlp~ ($r$ = `r cor.tlp`, $p$ = `r p.tlp`, Fig. \@ref(fig:erd-hyd)**b**), less negative $\mathrm{\Psi}$~88,stem~ and thus greater vulnerability to xylem embolism from cavitation ($r$ = `r cor.stem.88`, $p$ = `r p.stem.88`, Fig. \@ref(fig:erd-hyd)**c**), and narrower above-ground hydraulic safety margins ($\mathrm{\Psi}$~min~ - $\mathrm{\Psi}$~88,stem~, $r$ = `r cor.hsm`, $p$ = `r p.hsm`, Fig. \@ref(fig:erd-hyd)**d**).

## Effective rooting depths, mortality and hydrological droughts

Among the seven census-intervals over 1982-2015 for which we analysed relationship of ERD with mortality rates, six intervals were associated with occurrence of one or more El-Niño event [@Detto.2018; @Condit.2017]. Out of the six intervals with El-Niño events ERD explained 30-40% variation in mortality rates among evergreen species during five intervals--1982-1985, 1985-1990, 1995-2000, 2000-2005 and 2005-2010--such that species mortality rates decreased with deeper ERD ($p$-values < 0.05 for four intervals and 0.06 for one interval; Fig. \@ref(fig:erd-mort-census-evg)). For deciduous species ERD explained 11-16% variation in mortality in four intervals with El-Niño events, but p-values were not significant (Fig. S11).

\
Our analysis of hydrological droughts ranged from 1990 to 2015 and revealed distinctive extreme, prolonged hydrological droughts for the census-intervals 1995-2000, 2000-2005 and 2005-2010, for which ERD explained significant mortality (Fig. \@ref(fig:swp-droughts)). ERD also explained significant mortality in earlier droughts [1982-1985 and 1985-1990; @Condit.2017], not covered by our $\mathrm{\Psi}_{\textrm{soil}}$ estimates.

\
On an average, species exposure to water-stress (% days $\mathrm{\Psi}_{\textrm{soil},z=\textrm{ERD}}$ < $\mathrm{\Psi}$~crit~) exponentially declined with ERD, indicating that species with shallower ERD spent greater time under significant hydrological drought, and thus likely experienced greater hydraulic risk (also see SI). Exposure to water-stress increased over the three periods for which ERD explained significant mortality (Figs. \@ref(fig:erd-time-evg), S11), although it was also high in 1990-1995 for the shallowest ERD (Fig. \@ref(fig:erd-time-evg)), but without elevated mortality rates (Fig. \@ref(fig:erd-mort-census-evg)).

# Discussion

We introduce a novel approach for estimating ERD using 25 years of tree growth, species-specific leaf vulnerability curves, modeled soil water potential profiles, and observed VPD (Eq. \@ref(eq:M-chosen), Figs. \@ref(fig:schematic), \@ref(fig:sp-erd)). Our predictions of ERD were consistent with estimates using $\mathrm{\delta}$^2^H~xylem~ (Fig. \@ref(fig:erd-iso)). Our analyses suggest that co-occurring large canopy tree species with deeper ERD were associated with higher above-ground hydraulic efficiency, but lower safety (Fig. \@ref(fig:erd-hyd)). Nevertheless, evergreen but not deciduous species with deeper ERD showed significantly lower mortality rates than shallower ERD species (Fig. \@ref(fig:erd-mort-census-evg), S11). This ERD-mortality relationship was significant in five out of six census-intervals that had experienced one or more El-Niño events---over seven census-intervals studied in total (1982-2015). Quantifying extreme droughts in the whole soil column over 1990-2015 revealed that ERD explained significant levels of mortality in periods when extreme soil droughts prolonged the dry season water-stress (Fig. \@ref(fig:swp-droughts)). Species exposure to water-stress exponentially declined with deeper ERD (Fig. \@ref(fig:erd-time-evg)). Because extreme, sustained water-stress increases risk of mortality via hydraulic-failure and/or carbon starvation, deeper ERD may mitigate drought-induced mortality by limiting exposure to water-stress [@Rowland:2015; @McDowell:2018]. We thus demonstrate, for the first time, a link between species trade-offs in above- and below-ground hydraulic traits, drought exposure through extreme hydrological droughts quantified over the whole soil-column, and large ($\ge$ 10 cm DBH) tree mortality across several El-Niño events over 35 years.

## Drought strategies designed to mitigate realised hydraulic risks

If investments in stress-tolerance traits come at a cost, such insurance may not pay-off if the risk of stress is not realised. Our findings suggest that species with investment in deep roots can afford the hydraulically efficient, but risky suite of traits (Fig. \@ref(fig:erd-hyd)), because access to a reliable deep-water resource ensures that for them hydraulic risk is not realised (Fig. \@ref(fig:erd-time-evg)), whereas shallow rooted species pay the cost of hydraulic safety in terms of efficiency, adapted for an environment in which hydraulic risk is significant, as extreme droughts cause exponentially greater water-stress in shallow soil layers (Figs. Figs. \@ref(fig:erd-time-evg), S6).

\
Extreme, prolonged El-Niño droughts (Fig. \@ref(fig:swp-droughts)) in our study may have crossed even the greater tolerance limits of shallow-rooted species leading to their greater mortality compared to deep-rooted species (Fig. \@ref(fig:erd-mort-census-evg)). Hydraulic risk for shallow rooted species may have been exacerbated in our 35-yr study period in which extreme water-deficit years were more frequent than those in the last century [@Condit.2017]. Greater survival of deep-rooted species that we observed may not continue into the future if droughts intensify.

\
We found significant ERD-mortality relationships during extreme water-stress for evergreen but not deciduous species (Fig. \@ref(fig:erd-mort-census-evg), S11); consistent with the expectation that deciduous species can also avoid water-stress via leaf-drop. This is also consistent with the observation that species distributions along local (BCI) and regional (Panama) moisture gradients are correlated with $\mathrm{\Psi}$~tlp~ for evergreen but not deciduous species [@Kunert.2021]. Studies that simultaneously assess coordination between leaf phenology, rooting depth and hydraulic traits are almost absent for the tropics and warrant future consideration [@Oliviera.2021]. Our analyses for ERD-hydraulic traits relationships were limited to only seven species, but covered all species of different leaf phenologies (Fig. \@ref(fig:erd-hyd)), so here we assume that the trend in the ERD-hydraulic trait relationships holds true across all species of different leaf phenologies. We also found that the range of ERDs overlapped across leaf phenology, but evergreen species tend to have deeper ERDs (Fig. \@ref(fig:sp-erd)), consistent with observations by @Meinzer:1999 and elsewhere [@Smithmartin.2020; @Fan:2017; @Oliviera.2021]. Whether deciduous species also have more efficient and vulnerable hydraulics at BCI as is observed elsewhere remains to be studied [@Markesteijn:2010; @Markesteijn:2011; @Xu:2016; @Gleason.2016]. Leaf phenologies at BCI are numerous and complicated and warrant further research.

## Drought exposure integral to assessing drought-induced mortality

Our study brings attention to the need for assessing drought sensitivity in terms of species drought exposure and realized hydraulic risk by accounting for hydrological drought and trees' rooting depths. We found that species accessing deeper water had greater xylem vulnerability to embolism and narrower branch hydraulic safety margins (Fig. \@ref(fig:erd-hyd)). These traits are commonly identified as proxies for mortality risk [@Anderegg:2016], but in fact were associated with species with less drought exposure and had lower mortality. Hydraulic risk was balanced by investment in deep roots (Figs. \@ref(fig:swp-droughts), \@ref(fig:erd-mort-census-evg)).

\
Our results are consistent with recent studies that analysed rooting or water-sourcing depths versus hydraulic traits and mortality rates during extreme droughts [@Johnson:2018fad; @Nardini:2015; @Venturas.2016; also see, @Brum.2019; @Brum.2017; @Rowland:2015]. Globally, large trees tend to exhibit greater growth reductions, lower post-drought resilience, and greater increases in mortality relative to their understory counterparts [@Bennett:2015; @Phillips:2010]. Our finding that deep-water access buffers drought-induced mortality in large trees is relevant for understanding drought resistance, resilience and recovery [@Bennett:2015; @McGregor.2020]. Future studies should test the ERD-mortality relationship on a greater number of species.

\
Our result of lower mortality in deep-rooted trees contrasts with the inverse model finding of @Chitra-Tarak.2018 (hereafter, CT2018) in which deeper ERD species in a South-Asian seasonally dry tropical forest had higher mortality in a rare, prolonged drought. The hydrological model in CT2018 revealed that the multi-year drought exhausted the deep soil and even bedrock water availability (also see, @Goulden.2019; @Ivanov:2012). In contrast, our hydrological modeling at BCI finds that $\mathrm{\Psi}$~soil~ for depths deeper than 2.9 m did not cross $\mathrm{\Psi}$~crit~ of the most sensitive tree species (`r min.psi_crit` MPa) for all the dry seasons and droughts during 1990-2015 (see SI). Mean annual rainfall of 1095 mm in CT2018 compared to `r round(rain.man.stats$rain)` mm at BCI, and precipitation to potential evapotranspiration ratio of nearly one in CT2018 compared to nearly two in BCI, are major factors in the different mortality responses. At BCI deep soil layers were recharged annually (Fig. S6), whereas in CT2018 they were not. The contrast between the two studies highlights the combined role of seasonal precipitation input and site-specific hydrology in modulating mortality-risk of deep-rooted species.

## Modeling effective rooting depths at the tree community level

By estimating ERD for large trees of `r length(erd.sp)` canopy tree species (Fig. \@ref(fig:sp-erd)), we make an important advance in modeling effective rooting depths at the community level in species-rich tropical forests. Our best ERD model is a key improvement over the model of CT2018 as we employ a physiologically meaningful representation, fewer parameters and corroboration with tree and soil data for $\mathrm{\delta}$^2^H.

\
Our ERD model predicts daily maximum diurnal $\mathrm{\Psi}$~leaf~ assuming it is equivalent to $\mathrm{\Psi}$~soil~. Our model ignores other factors that buffer soil drying such as stem water storage capacitance [@Wolfe.2017], and may thus have overestimated ERD, especially in those species for which capacitance is important as a drought-avoidance strategy, for example, the deciduous species [@Borchert.2005jma]. Future studies should investigate the role of capacitance on estimating ERD.

\
While ERD models have the potential to estimate ERDs for entire tree communities, during their development phase, ERD models may need to be validated against data for a subset of representative species of the community, as this study did. ERD models should be tested across varied climates and forest types, covering contrasting plant strategies and possibly seasonality in ERDs. Direct observations of rooting depths and stable water isotopes based water-sourcing depths will be important datasets for such validation; although the interpretation of isotopic data is still under research [@Deurwaerder.2020; @Adams.2020; @Bowers.2020].

## Future directions

The relationships that we identify between above- and below-ground traits, vertical profile of soil water status and mortality rates are important for representing diversity in DGVMs, which intrinsically rely on the parameterization of contrasting life history strategies [@Scheiter:2013] and the simulation of competition between those strategies. In the context of trait filtering models, if we used the hydraulic trait information without knowledge of their relationship to rooting depth, models would likely kill the 'risky' strategy trees in droughts, which would in fact be the opposite result from that observed in this study. A key outcome of this study is thus the relationships between hydraulic traits and ERD that could be plugged into a DGVM of BCI. To assimilate ERDs, DGVMs could vary rooting parameters such that the centroid of the species water-uptake profiles match ERDs. We found that LMA and WSG were strong predictors of leaf hydraulic vulnerability curves (Fig. \@ref(fig:ab-fit)). Albeit future studies should undertake sensitivity analyses for uncertainties involved, our finding offers the promise of a greater ability to parameterize the 'hard' hydraulic traits with the abundant 'soft' trait data, thus allowing for a better representation of forest hydrodynamics. The relationships between ERD and above-ground hydraulic traits that we find thus provide important insights on how to model rooting depths and their coordination or trade-offs with other traits in order to better represent the functional diversity of tropical forests, and their trajectories into the future.

\
Our inverse ERD model was parameterized on 5-year growth data, with five data points over a 25-yr period, which decoupled climate events and demographic outcomes. Future studies could better constrain the ERD model with higher frequency growth data such as those from dendrometer bands. At high temporal resolution, however, the role of reversible dehydration in tree diameter change increases [@Chitra-Tarak.2016; @Chitra-Tarak.2015; @Mencuccini.2017], but that may provide an avenue to include stem water storage and dynamic rooting depths in ERD models. Three of our exploratory ERD models included LAI seasonality, but we did not select them as they worsened the fit with growth data for many species (Fig. S9). Our interpretation of this result is that VPD and leaf hydraulic vulnerability curves may be adequate to explain inter-census differences in growth (via stomatal control), but also acknowledge that our estimates of seasonality of LAI is a tentative estimate that combines leaf-fall data and the timing of leaf-gain backtracked from leaf lifetime, and omits inter-annual variation. Species-level leaf-fall data from litter-traps that we used includes large within-species variability in leaf-fall timing, and so may not have captured  tree-level seasonality in deciduousness (see SI). Future studies may improve models of LAI seasonality and make use of data from new technologies such as drone based monitoring of LAI in species-rich tropical forests [@Park.2019].

\
While our ERD model empirically predicts growth via estimating $\mathrm{\Psi}$~leaf~ from $\mathrm{\Psi}$~soil~ of a specific depth, mechanistic models that account for plant hydrodynamics and other processes influencing growth are likely to predict growth more accurately, and thus ERD and hydraulic risk [@Christoffersen:2016kd2d; @Sperry.1998; @Yang.2019; @Duursma:2018]. Although data needs for parameterizing such models could be greater (e.g. hydraulic vulnerability curves and capacitance for roots, stems and leaves), adding a degree of parameter uncertainty in community-wide application, such models hold greater promise in improving our understanding of plant physiology [e.g. see @Johnson:2018fad].

\
As in CT2018, with water-availability resolved for a 1-D column, we interpret our ERD estimates as revealing relative differences among species’ effective rooting depths rather than absolute depths.  To estimate absolute depths and topographic variation in ERD within or across species, future studies may use soil moisture dynamics from a distributed hydrological model [e.g. @Schwantes:2018]. We note that data availability on soil water retention curves ($\mathrm{\Psi}$~soil~ vs. VWC), hydraulic conductivity (K~soil~) by depth, soil moisture by depth, stream discharge and evapotranspiration were important for effective calibration of our 1-D hydrological model. We recommend widespread and coordinated collection of these variables as well as water-table levels in forest-inventory sites to allow for estimation of trees' water environments.

## Conclusions

Establishing relationships between environment, traits and demographic outcomes of plants is imperative for developing a predictive plant ecology. Trees' rooting depths and actual water environments through hydrological rather than meteorological droughts are nonetheless rarely studied. To our knowledge this is the first study to test for a mechanistic link between plant-available water in the whole-soil column, trees' above- and below-ground hydraulic architecture and long-term mortality outcomes for a species-rich forest. We report here that deep-water access plays a role in mitigating mortality of otherwise vulnerable stem hydraulics. This has important implications for our predictive understanding of tropical forest dynamics under current and future climate. Our community-scale framework for modeling effective rooting depths and leaf vulnerability curves indicates the possibilities in expanding the use of these critical, rare observations in species-rich forests towards community-scale generalizations.

# Acknowledgements

This research was supported as part of the Next Generation Ecosystem Experiments-Tropics, funded by the U.S. Department of Energy, Office of Science, Office of Biological and Environmental Research. Funding was also provided to RCT and SMM through an NSF grant (1137366). LR was deputed to Indo-French Cell for Water Sciences on an IRD Fellowship. An NSF grant to Smithsonian Institution’s ForestGEO (Dimensions 1046113) partially supported this work via RCT and SMM. BTW was supported by the National Institute of Food and Agriculture, U.S. Department of Agriculture, McIntire Stennis project under LAB94493. LS was supported by National Science Foundation award 2017949. RF was supported by the National Center for Atmospheric Research, which is funded by the National Science Foundation. The BCI forest dynamics research project, founded by Stephen P. Hubbell and Robin B. Foster and sustained for many years by Richard Condit, is now managed by Rolando Perez, Suzanne Lao and Stuart Davies under the ForestGEO program of the Smithsonian Tropical Research in Panama. Numerous organizations have provided funding, principally the US National Science Foundation, and hundreds of field workers have contributed.

# Author contribution

Conceptualization: RCT. Data curation: BF, SP for climate drivers. SP for stream discharge. MTT for GPP, ET, TDR, bulk density. SK for manual GWC and GWC vs $\mathrm{\Psi}_\textrm{soil}$. NK, JZ, KAT, LS for leaf hydraulics, BW for stem hydraulics. RP and SA for the BCI 50ha plot censuses. JW for maximum tree height, WSG, LMA, deciduousness, leaf fall and leaf lifetime. Formal Analysis: RCT, CX. MTT for upgrading GPP-VPD relationship and LAI seasonality. Funding acquisition: SMM, JC, CX, NM, LK. Investigation: RCT, CX (SMM and LR for an exploratory version). Methodology: RCT, CX (SMM and LR for an exploratory version). Project administration: RCT, Resources: SMM, BF, SP, MTT, SK, NK, JZ, KAT, LS, BW, JW. Software: RCT; RF, RK, CF, CX, RCT for ELM-FATES. Supervision: CX, SMM, BN, Validation: RCT. Visualization: RCT, Writing – original draft: RCT. Writing – review & editing: All authors. See https://casrai.org/credit/ for the taxonomy of credits.

# Data Availability

ELM-FATES source code, simulation outputs for best-fit parameter ensemble members and all the R scripts to reproduce the manuscript are available [@Chitra-Tarak.2020]. Data sources for all other datasets used are provided throughout the manuscript.

# References
<div id="refs"></div>

# Figure Captions 
\
Figure \@ref(fig:schematic): A schematic outlining the methods workflow, which combines hydrology, physiology and demography of the tropical forest at Barro Colorado Island to inform the inverse model for rooting depths; validates it, and tests hypotheses pertaining to relationship of rooting depths with above-ground hydraulic traits, drought exposure, and drought-induced mortality.

\
Figure \@ref(fig:hydro-calib): ELM-FATES calibration and evaluation. Observations (red lines and points) versus simulations (gray lines) from 100 best-fit ensemble-member runs of ELM-FATES are shown for daily Volumetric Water Content (VWC) by depth for horizontal time domain reflectometry (TDR) probes at three depths (0.1, 0.4 and 1 m; **a**) and for an average of three vertical TDR probes (0-0.15 m; **b**); monthly stream discharge (Discharge, **c**) and monthly evapotranspiration from the flux tower (ET, **d**). Red points in panel **a** show average, manual plot-wide observations of VWC with 95% CI (error bars). All observations (red lines and points) were in-sample, except for VWC data from the horizontal probes (red lines in **a**) which were out-of-sample. Values in inset are average RMSE across the 100 best-fit simulations. For panel **a**, values at the top are for manual, plot-wide VWC and those at the bottom are for VWC from TDR probes. VWC is in units of cm^3^ cm^-3^, while ET and Discharge are in mm month^-1^.

\
Figure \@ref(fig:swp-droughts): Occurrence of extreme hydrological droughts by census intervals (horizontal panels) at three representative depths: 0.1, 0.6 and 1 m (vertical panels). Mean (black lines) and lower half of 95% distribution (grey areas) of soil water potential $\mathrm{\Psi}_{\textrm{soil},z}$ for a given DOY and depth $z$ are shown (same across all census intervals). $\mathrm{\Psi}_{\textrm{soil},z}$ by DOY for a year (colored line) is only shown if at least one day of the year $\mathrm{\Psi}_{\textrm{soil},z}$ was more negative than the 5^th^ percentile of $\mathrm{\Psi}_{\textrm{soil},z}$ for that day of year (DOY). Note the distinctive features of extreme droughts that occurred during the three periods indicated by a $*$: all three periods featured extreme hydrological droughts that either prolonged over the dry season (1995-2000), or occurred in the wet season, effectively extending the dry season (2000-2005, 2005-2010).

\
Figure \@ref(fig:ab-fit): Parameters $A$ (panel **a**) & $B$ (panel **b**) for the leaf vulnerability curves that were fitted to observed data on $K_\textrm{leaf}$ vs. $\mathrm{\Psi}$~leaf~ for `r length(sp.exp.param$sp)` species (Eq. \@ref(eq:vc-exp)) vs. those that were predicted from a set of models based on trait-proxies (Eqs. \@ref(eq:a-mod)-\@ref(eq:b-mod)). Goodness-of-fit ($R$^2^) and significance levels are given in inset.

\
Figure \@ref(fig:erd-iso): Modeled effective rooting depth (ERD; mean $\pm$ 1SE; m) versus dry-season stable isotopic concentration $\mathrm{\delta}$^2^H~xylem~ (mean $\pm$ 1SE; ‰) for `r length(unique(erd.iso$sp))` canopy species from Barro Colorado Island. $\mathrm{\delta}$^2^H~xylem~ data are from @Meinzer:1999.

\
Figure \@ref(fig:sp-erd): Modeled effective rooting depth (ERD; mean $\pm$ 1SE in m) for `r  length(erd.sp)` large, canopy species of Barro Colorado Island. A species ERD is defined as the median, across 100 hydrological realizations, of soil layer depth $z$ at which soil water dynamics ($\mathrm{\Psi}_{\textrm{soil,} z}$) best explained observed dynamics of species growth for each realization. As soil layers in the hydrological model are discretely resolved into exponentially increasing depths, so is the ERD axis. Species are color coded by leaf habit.

\
Figure \@ref(fig:erd-hyd): Modeled effective rooting depth (ERD) versus hydraulic properties for seven canopy species found on Barro Colorado Island (Panama). Spearman's $r$ and significance levels are given in panel insets. Linear model fits (blue lines) with confidence bands (gray area) are shown for significant fits at $\alpha$ = 0.05. See Fig. S10 for the full correlation matrix. Species are color coded by leaf habit. Multiple species in Facultative Deciduous leaf-habit are distinguished by shapes.

\
Figure \@ref(fig:erd-mort-census-evg): Mortality rate (mean $\pm$ 1SE in % yr^-1^) versus modeled effective rooting depth (ERD; mean $\pm$ 1SE in m) for `r nrow(mrate.depth.mean %>% subset(deciduous == "E"))` evergreen, canopy species found on Barro Colorado Island over seven census intervals (1981-2015). $R$^2^ and significance levels for linear model fits are given in panel insets. Model fits (blue lines) with confidence bands (gray area) are only shown for periods with significant fits. Census intervals, for which fits are significant at $\alpha$ = 0.05 are identified by superscript $**$, while those significant at $\alpha$ = 0.1 by superscript $*$.

\
Figure \@ref(fig:erd-time-evg): Modeled effective rooting depth (ERD; horizontal-axis) versus time spent beyond critical hydraulic threshold (vertical axis) by census interval (colored bars) for `r nrow(mrate.depth.mean %>% subset(deciduous == "E"))` evergreen species included in the mortality analyses (Fig. \@ref(fig:erd-mort-census-evg)). Each bar represents the average time species of the same ERD spent beyond species-specific critical hydraulic thresholds in a given interval, that is, the proportion of days for which $\mathrm{\Psi}_{\textrm{soil},z=\textrm{ERD}}$ was more negative than species $\mathrm{\Psi}$~crit~, defined as $\mathrm{\Psi}$~20,leaf~, and where $z$ is the soil depth matching species ERD. Standard error of the mean are shown over each bar when available. Note the squared y-axis scale.

# Figures

```{r schematic, eval = TRUE, echo = FALSE, message = FALSE, fig.align = 'center', out.width = "80%", fig.pos = 'H', results = "asis", fig.cap = paste0("\\label{lab-schematic}")}
knitr::include_graphics(file.path("schematic_easy/schematic_easy.001.jpeg"))
```

```{r hydro-calib, eval = TRUE, echo = FALSE, message = FALSE, fig.align ='center', out.width = "80%", fig.pos = 'H', results = "asis", fig.cap = paste0("\\label{lab-hydro-calib}")}
knitr::include_graphics(file.path("figures/2019-10-14_5000/best-fits/hydro-calib2.jpeg"))
```

```{r swp-droughts, eval = TRUE, echo = FALSE, message = FALSE, fig.align ='center', out.width = "100%", fig.pos = 'H', results = "asis", fig.cap = "\\label{lab-swp-droughts}"}
knitr::include_graphics(file.path("figures/PhenoDemoTraitsPsi/psi_model_daily_bestfit_params.top.few_CI_full_interval_panels_climatology_over_study_period_q5_by_depth.jpeg"))
```

```{r ab-fit, eval = TRUE, echo = FALSE, message = FALSE, fig.align ='center', dpi = 96, out.width = "60%", fig.pos = 'H', results = "asis", fig.cap = paste0("\\label{lab-ab-fit}")}
knitr::include_graphics(file.path("figures/PhenoDemoTraitsPsi/kmax_by_psi/Leaf/A_B_data_vs_model.jpeg"))
```

```{r erd-iso, eval = TRUE, echo = FALSE, message = FALSE, fig.align = 'center', out.width = "80%", fig.pos = 'H', results = "asis", fig.cap = paste0("\\label{lab-erd-iso}")}
knitr::include_graphics(file.path("figures/PhenoDemoTraitsPsi/psi.corr_best.depth_xylem_sap_deltaD_phenology_Meinzer_gr.Psi.VPD.jpeg"))
```

```{r sp-erd, eval = TRUE, echo = FALSE, message = FALSE, fig.align ='center', out.width = "60%", fig.pos = 'H', results = "asis", fig.cap= paste0("\\label{lab-sp-erd}")}
knitr::include_graphics(file.path("figures/PhenoDemoTraitsPsi/ERD_by_sp_large_canopy.jpeg"))
```

```{r erd-hyd, eval = TRUE, echo = FALSE, message = FALSE, fig.align ='center', out.width = "60%", fig.pos = 'H', results = "asis", fig.cap = paste0("\\label{lab-erd-hyd}")}
knitr::include_graphics(file.path("figures/PhenoDemoTraitsPsi/erd.stem.traits.jpeg"))
```

```{r erd-mort-census-evg, eval = TRUE, echo = FALSE, message = FALSE, fig.align ='center', out.width = "100%", fig.pos = 'H', results = "asis", fig.cap = paste0("\\label{lab-erd-mort-evg}")}
knitr::include_graphics(file.path("figures/PhenoDemoTraitsPsi/mortality_by rdi.gr_evergreen.jpeg"))
```

```{r erd-time-evg, eval = TRUE, echo = FALSE, message = FALSE, fig.align ='center', out.width = "40%", fig.pos = 'H', results = "asis", fig.cap = paste0("\\label{lab-erd-time-evg}")}
knitr::include_graphics(file.path("figures/PhenoDemoTraitsPsi/mfac vs. rdi.gr_evergreen.jpeg"))
```

# Tables
```{r symbols, results = 'asis', echo = FALSE}
knitr::kable(symbols.table, caption = 'The symbol, definition and units of key traits used in the simulations and analyses.', booktabs = TRUE, escape = FALSE)
```

#
```{r hypo, results = 'asis', echo = FALSE}
knitr::kable(hypo.table, caption = 'Hypotheses for association between Effective Rooting Depth (ERD) and above-ground hydraulic traits.', booktabs = TRUE, escape = FALSE, col.names = gsub("[_]", "-", gsub("[.]", " ", names(hypo.table))))
```